---
title: "Claessen Project"
output: 
  html_notebook:
    toc: true
    toc_depth: 3
    toc_float: 
      collapsed: false
      smooth_scroll: true
    theme: cosmo
    df_print: paged
    highlight: tango
    # code_folding: hide
---

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

# Load libraries

```{r}
library(bvpSolve)
library(ggplot2)
library(plyr)
library(gplots)
options(scipen = 999)
```

### Add info from excel:

```{r}
#my_data=c("4632","3776","3661","5383","2619","3617","3655","3061","3048","4390","2093","3039","35","42","32","34","37","39","80","97","111","91","89","64","21","26","26","22","33","27","14.1","13.3","14.4","15.4","15.9","16.5","25.99","16.85","21.29","20.62","11.73","17.28")
my_data <-  c("4632","3655","35","80","21","14.1","25.99","96","23.3","3776","3061","42","97","26","13.3","16.85","98.3","41.7","3661","3048","32","111","26","14.4","21.29","98.7","39.2","5383","4390","34","91","22","15.4","20.62","98.1","29.9","2619","2093","37","89","33","15.9","11.73","97.8","66.6","3617","3039","39","64","27","16.5","17.28","93.1","43.4","4287","5239","35","125","51","16.3","24.46","98.4","35","3479","4631","39","99","27","14.5","26.89","97.8","41","3968","4284","38","104","30","14.1","24.31","98.1","38.8","4745","5857","37","84","24","15.4","28.65","97","28.7","3364","3956","35","86","25","15.4","21.22","97.2","34.7","2091","2767","39","82","26","14.6","12.77","96.6","37.6","3544","4797","39","85","25","15.3","29.24","97.4","40.1","3733","4480","36","105","38","15.7","24.71","98.1","55.8","1338","1456","29","53","21","18.2","8.44","91.2","40","1677","1395","30","67","25","15.3","10.46","92.6","33.1","800","755","28","90","16","13.3","5.48","98","17.3","1128","1130","34","61","26","17.1","9.24","89.2","37.6","1370","1365","23","43","21","15.8","13.55","86.8","45.3","816","859","23","51","19","15.1","5.36","91","35.5","1332","1129","21","60","18","15.7","10.8","91.8","22.2")
my_data <- as.data.frame(matrix(data=as.numeric(my_data),ncol=9,byrow=T))
rownames(my_data) <-c(paste("C",seq(14),sep=""),paste("CTEPH",seq(7),sep=""))
colnames(my_data) <- c("VO2","VCO2","PaCO2","PaO2","PvO2","Hb","Q_exp","satO2_a","satO2_v")
my_data
```
PATIENT samples

```{r echo=FALSE}
# my_data_PAT<-c("1338","1456","29","53","21","18.2","8.44","1677","1395","30","67","25","15.3","10.46","800","755","28", "90","16","13.3","5.48","1128","1130","34","61","26","17.1","9.24","1370","1365","23","43","21","15.8",  "13.55","816","859","23","51","19","15.1","5.36","1332","1129","21","60","18","15.7","10.8")
# my_data_PAT <- as.data.frame(matrix(data=as.numeric(my_data_PAT),ncol=7,nrow=7,byrow=TRUE))
# rownames(my_data_PAT) <-paste("CTEPH",seq(7),sep="")
# colnames(my_data_PAT) <- c("VO2","VCO2","PaCO2","PaO2","PvO2","Hb","Q")
# my_data <- rbind(my_data,my_data_PAT)
#my_data
```

EXTRA 2 samples from the paper (average values for the healthy and HFpEF patients)
NO INFO FOR PaCO2 in the paper --> i will put 35 / 30

```{r}
paper_HEALTHY <- c("1620","1940","35","97","21","14.1","14","100","34")
paper_HFPEF <- c("1090","1260","30","96","21","13.4","10.2","100","34")
my_data_PAPER <- c(paper_HEALTHY,paper_HFPEF)
my_data_PAPER <- as.data.frame(matrix(data=as.numeric(my_data_PAPER),ncol=9,nrow=2,byrow=TRUE))
colnames(my_data_PAPER) <- colnames(my_data)
rownames(my_data_PAPER) <-c("REFCTR","REFPAT")
my_data<-rbind(my_data,my_data_PAPER)
my_data
```

## Application of Algorithm 1:

O2 pathway parameter estimation from an individual's exercise measurements.

1. Input:

   a) CPET-measurements at peak exercise for patient i: `VO2, VCO2, PaCO2, PaO2, PvO2, Hb`
   b) Constants: `k, PlO2, p50, RCtlmito, RHFmito`
  
2. Solve O2 transport equations (1)-(2), (7)-(11) with boundary conditions (3)-(6)
3. Output:

   a) Patient i O pathway parameters at peak exercise: `Q, DL, DM, VA, vmax`
   b) Patient i O2 tension in the alveolus, mitochondria, pulmonary and muscle capillaries: `PAO2, Pmito, PLcap(t), PMcap(t)`

### Define Constants:

* k= 1.159 (mL O2 mL-1 air mmHg -1) --> units conversion constant
* PIO2= 150 mmHG (from http://www.partone.litfl.com/oxygen_cascade.html)--> Inspired partial pressure of oxygen
* p50=0.24 --> Mitochondrial oxygen affinity (mmHg)
* RCtlmito= 1.8 --> reserve mitochondrial respiration capacity, for HEALTHY patients, derived from Esposito et al 11
* RHFmito= 2.0 --> sameas above
* T (capillary transit time)= 0.5s or 0.00833 min --> from papers from Guidos email...not sure but will try with this value first

```{r}
ct_k=1.159
ct_piO2=150
ct_p50=0.24
ct_Rctlmito=1.8
ct_RHFmito=2
ct_T=0.00833
```

##### First, we will calculate Q for all the samples and then plot the calculated vs the measured Q:
Parameters needed:

* VO2 --> my_data[1]
* PaO2 --> my_data[4]
* Hb --> my_data[6]
* PvO2 --> my_data[5]
* SatO2_v --> my_data[9] 
* SatO2_a --> my_data[8]
```{r}
calcQ <- function(p1){
  res=p1[1]/((0.0032*p1[4]+1.4*p1[6]*p1[8])-(0.0032*p1[5]+1.4*p1[6]*p1[9]))*10;
  return(res)
}
calcQ(my_data[1,])
```
```{r}
round(apply(my_data,1,calcQ),2)
my_data$Q_calc <- round(apply(my_data,1,calcQ),2)
```

Plot correlation:

```{r}
my_data$group <- c(rep('control',14),rep('patient',7),rep('ref',2))
```


```{r echo =F}
#ggscatter(my_data,x='Q_exp',y='Q_calc',add="reg.line") +
# stat_cor(aes(label=paste(..rr.label..,..p.label..,sep="~`,`~")),label.x=3)
```

```{r echo=F}
#specific intersect
modelINT <- lm(I(Q_calc)~0+my_data$Q_exp,data=my_data)
modelINT_cor <- cor(my_data$Q_calc,my_data$Q_exp)**2
coeffINT=coefficients(modelINT)
eqINT = paste0("y = ", round(coeffINT[1],2), "*x", ", R2 = ", round(modelINT_cor,2))
#q_plot + geom_point(aes(color=group)) +
#  geom_abline(intercept=0,slope=coeffINT[1]) +
#  ggtitle(eqINT)
```

```{r}
plot_data <- my_data[-5,]
model <- lm(Q_calc~Q_exp,data=plot_data)
model_cor <- cor(plot_data$Q_calc,plot_data$Q_exp)**2
coeff=coefficients(model)
eq = paste0("y = ", round(coeff[2],2), "*x + " ,round(coeff[1],2))
eq_cor=bquote(R^2 ~"=" ~ .(round(model_cor,2)))
#eq_cor=expression(paste("R"^"2"))
q_plot <- ggplot(plot_data, aes(Q_exp, Q_calc ))
q_plot + geom_point(aes(color=group)) +
  #geom_abline(intercept=coeff[1],slope=coeff[2]) +
  geom_smooth(method=lm,se=F,color="black") +
  #ggtitle(eq) +
  annotate(geom="text", x=5, y=40, label=eq, color="black",hjust=0) +
  annotate(geom="text", x=5, y=38, label=eq_cor, color="black",hjust=0) +
  ggtitle("Correlation between Calculated and Experimental Q") +
  ylab ("Calculated Q (L/min)") +
  xlab ("Experimental Q (L/min)")
  #geom_text(x=7, y=30, label=eq_cor)
```

### Calculate other variables and parameters for the final ODE

```{r}
calcVA <- function (input){
  res=as.numeric(input[2])/(ct_k*as.numeric(input[3]))
  return (res)
}
my_data$VA <- round(apply(my_data,1,calcVA),2)
```

```{r}
calcPAO2 <- function(input) {
  res=ct_piO2-(as.numeric(input[1])/(as.numeric(input[12])*ct_k))
  return(res)
}
my_data$PAO2 <- round(apply(my_data,1,calcPAO2),2)
```

```{r}
calcVMAX <- function(input,rmito) {
  res=rmito*as.numeric(input[1])
  return(res)
}
#my_data$Vmax <- apply(my_data,1,calcVMAX,rmito=ct_Rctlmito)
vmax_ctl <- round(apply(my_data,1,calcVMAX,rmito=ct_Rctlmito),2)
vmax_pat <- round(apply(my_data,1,calcVMAX,rmito=ct_RHFmito),2)
my_data$Vmax <- c(vmax_ctl[1:14],vmax_pat[15:21],vmax_ctl[22],vmax_pat[23])
```

```{r}
calcPMITO <- function(input) {
  res=ct_p50/((as.numeric(input[14])/as.numeric(input[1]))-1)
  return(res)
}
# my_data$Pmito <- formatC(apply(my_data,1,calcPMITO), format = "e", digits = 2)
my_data$Pmito <- apply(my_data,1,calcPMITO)
```

### Summary of all the calculated variables

```{r}
my_data
```

### Plots as in the paper
Q vs VO2

```{r}
create_cor_plot <- function(pData, xVar, yVar,legx,legy,myxlim=c(0,100),myylim=c(0,100),axis=c("title","xlab","ylab")) {
  model <- lm(formula(paste(yVar,"~",xVar)),data=pData)
  model_cor <- cor(pData[,yVar],pData[,xVar])**2
  coeff=coefficients(model)
  eq = paste0("y = ", round(coeff[2],4), "*x + " ,round(coeff[1],2))
  eq_cor=bquote(R^2 ~"=" ~ .(round(model_cor,2)))
  q_plot <- ggplot(pData, aes_string(x=xVar, y=yVar ))
  q_plot + geom_point(aes(color=group)) +
    geom_smooth(method=lm,se=T,color="black") +
    annotate(geom="text", x=legx, y=legy, label=eq, color="black",hjust=0) +
    annotate(geom="text", x=legx, y=legy-(0.05*legy), label=eq_cor, color="black",hjust=0) +
    xlim(myxlim)+
    ylim(myylim) +
    ggtitle(axis[1]) +
    xlab (axis[2]) +
    ylab (axis[3])
}
```

Calculated Q

```{r}
# plot_data <- my_data[-5,]
plot_data <- my_data
plot_data$VO2 <- plot_data$VO2/1000
pTitle <- expression("Correlation between Calculated Q and V"["O"[2]])
pxLab <- expression("V"["O"[2]]*" (L/min)")
pyLab <- "Calculated Q (L/min)"
create_cor_plot(plot_data,"VO2","Q_calc",1,40,c(0,6),c(0,50),c(pTitle,pxLab,pyLab))
```

Experimental Q

```{r}
pTitle <- expression("Correlation between Experimental Q and V"["O"[2]])
pxLab <- expression("V"["O"[2]]*" (L/min)")
pyLab <- "Experimental Q (L/min)"
create_cor_plot(plot_data,"VO2","Q_exp",1,30,c(0,6),c(0,35),c(pTitle,pxLab,pyLab))
```

VA vs VO2

```{r}
pTitle <- expression("Correlation between V"["A"]*" and V"["O"[2]])
pxLab <- expression("V"["O"[2]]*" (L/min)")
pyLab <- expression("V"["A"]* " (L/min)")
create_cor_plot(plot_data,"VO2","VA",1,130,c(0,6),c(0,150),c(pTitle,pxLab,pyLab))
```

### Missing Plots

* Q vs DAV_O2 (DAV_O2 not calculated)

### Functions to Integrate the ODE and get the last parameters

#### DL and PLcap

* VO2 --> my_data[1]
* PaO2 --> my_data[4]
* Hb --> my_data[6]
* PvO2 --> my_data[5]
* SatO2_v --> my_data[9] 
* SatO2_a --> my_data[8]
* Q --> my_data[10]
* Pmito--> my_data[15]
* T (capillary transit time)= 0.5s or 0.00833 min --> from papers from Guidos email...not sure but will try with this value first


```{r}
# ct_T=0.00833
# ct_T=10
f_PL <- function(t,y,DL,params){
  # dy <- DL / (y[10]*ct_T*(0.0032*y[1]+1.4*y[6]*y[9])) * (y[13]-y[1])
  dy <- DL / (as.numeric(params[2])*ct_T*(0.0032*y[1]+1.4*as.numeric(params[1])*as.numeric(params[3]))) * (as.numeric(params[4])-y[1])
  list(dy)
}
```


```{r include=F}
t <- seq(0,ct_T,ct_T/20)
calc_PL <- function (data) {
  hb <- data[6]
  q <- data[10]
  satv <- as.numeric(data[8])/100
  pao2 <- data[13]
  mysol <- bvpshoot(yini=as.numeric(data[5]),yend=as.numeric(data[4]),x=t,func=f_PL,extra=20,maxiter=100000000,params=c(hb,q,satv,pao2))
  # mysol <- bvpshoot(yini=21,yend=64,x=t,func=f_PL,extra=20,maxiter=100000000,params=c(hb,q,satv,pao2))
  DL <- attr(mysol,"roots")
  plot(mysol[,1],mysol[,2],type="l")
  return(DL$root/10)
}
```

```{r}
options(scipen = 999)
par(mfrow=c(4,6))
plot_data$DL <- round(apply(my_data,1,calc_PL),2)
#dev.off()
```


```{r}
pTitle <- expression("Correlation between D"["L"]*" and V"["O"[2]])
pxLab <- expression("V"["O"[2]]*" (L/min)")
pyLab <- expression("D"["L"]* " (mL/min" %.% "mmHg)")
create_cor_plot(plot_data[c(-7,-2,-3),],"VO2","DL",0,150,c(0,6),c(0,170),c(pTitle,pxLab,pyLab))
```

#### DM and PMcap

```{r}
f_PM <- function(t,y,DM,params){
  dy <- - DM / (as.numeric(params[2])*ct_T*(0.0032*y[1]+1.4*as.numeric(params[1])*as.numeric(params[3]))) * (y[1]-as.numeric(params[4]))
  list(dy)
}
```


```{r include=F}
t <- seq(0,ct_T,0.0005)
calc_PM <- function (data) {
  hb <- data[6]
  q <- data[10]
  satv <- as.numeric(data[8])/100
  pmito <- data[15]
  mysol <- bvpshoot(yini=as.numeric(data[4]),yend=as.numeric(data[5]),x=t,func=f_PM,extra=50,maxiter=100000000,params=c(hb,q,satv,pmito))
  DM <- attr(mysol,"roots")
  plot(mysol[,1],mysol[,2],type="l")
  return(DM$root/10)
}
```

```{r}
par(mfrow=c(4,6))
plot_data$DM <- round(apply(my_data,1,calc_PM),2)
# dev.off()
```


```{r}
pTitle <- expression("Correlation between D"["M"]*" and V"["O"[2]])
pxLab <- expression("V"["O"[2]]*" (L/min)")
pyLab <- expression("D"["M"]* " (mL/min" %.% "mmHg)")
create_cor_plot(plot_data,"VO2","DM",0,100,c(0,6),c(0,120),c(pTitle,pxLab,pyLab))
```

## Extra Plots

### Histogram DM with respect to different groups (Fig 2D in the paper)
```{r}
# from http://www.sthda.com/english/wiki/ggplot2-barplots-quick-start-guide-r-software-and-data-visualization
#+++++++++++++++++++++++++
# Function to calculate the mean and the standard deviation
  # for each group
#+++++++++++++++++++++++++
# data : a data frame
# varname : the name of a column containing the variable
  #to be summariezed
# groupnames : vector of column names to be used as
  # grouping variables
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}
```

```{r}
df3 <- data_summary(plot_data, varname="DM", groupnames="group")
pTitle <- expression("Barplot of the different groups with respect to D"["M"])
pxLab <- "Groups"
pyLab <- expression("D"["M"]* " (mL/min" %.% "mmHg)")
ggplot(df3,aes(x=group,y=DM,fill=group)) + geom_bar(stat = "identity",position = position_dodge()) + geom_errorbar(aes(ymin=DM-sd, ymax=DM+sd),width=0.2,position=position_dodge(.9)) +
  ggtitle(pTitle) +
  xlab (pxLab) +
  ylab (pyLab)
```
### Same with all the variables

```{r}
units<-c("L/min","mL/min","mmHg","mmHg","mmHg","g/dL","L/min","%","%","L/min","L/min","mmHg","mL/min","mmHg","mL/min x mmHg","mL/min x mmHg")
units <- as.data.frame(matrix(units,nrow=1))
colnames(units) <- colnames(plot_data[,-11])
for (i in colnames(plot_data[,-11])){
  pd <- plot_data
  if (i == 'DL') {pd <- plot_data[-7,]}
  if (i =='Pmito') {pd$Pmito <- as.numeric(pd$Pmito)}
  df3 <- data_summary(pd, varname=i, groupnames="group")
  pTitle <- paste("Barplot of the different groups with respect to",i)
  pxLab <- "Groups"
  pyLab <- paste0(i," (",as.character(getElement(units,i)),")")
  p <- ggplot(df3,aes_string(x="group",y=i,fill="group")) + 
    geom_bar(stat = "identity",position = position_dodge()) +
    geom_errorbar(aes_string(ymin=paste(i,"-sd"), ymax=paste(i,"+sd")),width=0.2,position=position_dodge(.9))+
    ggtitle(pTitle) +
    xlab (pxLab) +
    ylab (pyLab)
  plot(p)
}
```
### Summary of all the data

```{r}
plot_data
```

# Repeat everything predicting VO2 instead

```{r}
my_data$DL <- plot_data$DL
my_data$DM <- plot_data$DM
my_data
```

```{r}
my_data_VO2 <- my_data[,c(1:11)]
my_data_VO2$Q_calc <- my_data_VO2$Q_exp
my_data_VO2
```


```{r}
calcVO2 <- function(p1){
  p1 <- as.numeric(p1[c(1:9)])
  res=p1[7]*((0.0032*p1[4]+1.4*p1[6]*p1[8])-(0.0032*p1[5]+1.4*p1[6]*p1[9]))/10
  return(res)
}
calcVO2(my_data_VO2[1,])
```
```{r}
round(apply(my_data_VO2,1,calcVO2),2)
my_data_VO2$VO2 <- round(apply(my_data,1,calcVO2),2)
kk <- my_data[,c(1,2,11)]
colnames(kk) <- c("VO2exp","VO2calc","group")
kk$VO2calc <-round(apply(my_data,1,calcVO2),2)
```
```{r}
pTitle <- expression("Correlation between Calculated and Experimental V"["O"[2]])
pxLab <- expression("Calc V"["O"[2]]*" (mL/min)")
pyLab <- expression("Exp V"["O"[2]]*" (mL/min)")
create_cor_plot(kk,"VO2calc","VO2exp",500,4500,c(0,5500),c(0,5500),c(pTitle,pxLab,pyLab))
```

```{r}
pTitle <- expression("Correlation between Experimental Q and Calculated V"["O"[2]])
pxLab <- expression("Calc V"["O"[2]]*" (mL/min)")
pyLab <- "Exp Q (L/min)"
create_cor_plot(my_data_VO2,"VO2","Q_exp",1,40,c(0,6000),c(0,50),c(pTitle,pxLab,pyLab))

pTitle <- expression("Correlation between Experimental Q and Experimental V"["O"[2]])
pxLab <- expression("Exp V"["O"[2]]*" (mL/min)")
pyLab <- "Exp Q (L/min)"
create_cor_plot(my_data,"VO2","Q_exp",1,40,c(0,6000),c(0,50),c(pTitle,pxLab,pyLab))

pTitle <- expression("Correlation between Calc Q and Exp V"["O"[2]])
pxLab <- expression("Exp V"["O"[2]]*" (L/min)")
pyLab <- "Calc Q (L/min)"
create_cor_plot(my_data,"VO2","Q_calc",1,40,c(0,6000),c(0,50),c(pTitle,pxLab,pyLab))
```
```{r}
my_data_VO2$VA <- round(apply(my_data_VO2,1,calcVA),2)
my_data_VO2$PAO2 <- round(apply(my_data_VO2,1,calcPAO2),2)
vmax_ctl <- round(apply(my_data_VO2,1,calcVMAX,rmito=ct_Rctlmito),2)
vmax_pat <- round(apply(my_data_VO2,1,calcVMAX,rmito=ct_RHFmito),2)
my_data_VO2$Vmax <- c(vmax_ctl[1:14],vmax_pat[15:21],vmax_ctl[22],vmax_pat[23])
my_data_VO2$Pmito <- round(apply(my_data_VO2,1,calcPMITO),2)
```
```{r}
pTitle <- expression("Correlation between V"["A"]*" and Calc V"["O"[2]])
pxLab <- expression("Calc. V"["O"[2]]*" (L/min)")
pyLab <- expression("V"["A"]* " (L/min)")
create_cor_plot(my_data_VO2,"VO2","VA",1,130,c(0,6000),c(0,150),c(pTitle,pxLab,pyLab))
```
```{r}
par(mfrow=c(4,6))
my_data_VO2$DL <- round(apply(my_data_VO2,1,calc_PL),2)
```

```{r}
pTitle <- expression("Correlation between D"["L"]*" and V"["O"[2]])
pxLab <- expression("V"["O"[2]]*" (mL/min)")
pyLab <- expression("D"["L"]* " (mL/min" %.% "mmHg)")
create_cor_plot(my_data_VO2[-7,],"VO2","DL",0,100,c(0,6000),c(0,120),c(pTitle,pxLab,pyLab))
```
```{r}
par(mfrow=c(4,6))
my_data_VO2$DM <- round(apply(my_data_VO2,1,calc_PM),2)
```

```{r}
pTitle <- expression("Correlation between D"["M"]*" and V"["O"[2]])
pxLab <- expression("V"["O"[2]]*" (L/min)")
pyLab <- expression("D"["M"]* " (mL/min" %.% "mmHg)")
create_cor_plot(my_data_VO2,"VO2","DM",0,100,c(0,6000),c(0,120),c(pTitle,pxLab,pyLab))
```


```{r}
units<-c("L/min","mL/min","mmHg","mmHg","mmHg","g/dL","L/min","%","%","L/min","L/min","mmHg","mL/min","mmHg","mL/min x mmHg","mL/min x mmHg")
units <- as.data.frame(matrix(units,nrow=1))
plots <- NULL
colnames(units) <- colnames(my_data_VO2[,-11])
for (i in colnames(my_data_VO2[,-11])){
  pd <- my_data_VO2
  if (i == 'DL') {pd <- my_data_VO2[-7,]}
  if (i =='Pmito') {pd$Pmito <- as.numeric(pd$Pmito)}
  df3 <- data_summary(pd, varname=i, groupnames="group")
  pTitle <- paste("Barplot of the different groups with respect to",i)
  pxLab <- "Groups"
  pyLab <- paste0(i," (",as.character(getElement(units,i)),")")
  p <- ggplot(df3,aes_string(x="group",y=i,fill="group")) + 
    geom_bar(stat = "identity",position = position_dodge()) +
    geom_errorbar(aes_string(ymin=paste(i,"-sd"), ymax=paste(i,"+sd")),width=0.2,position=position_dodge(.9))+
    ggtitle(pTitle) +
    xlab (pxLab) +
    ylab (pyLab)
  plot(p)
}
```
```{r}
# heatmap.2(my_data_VO2[,-11], key=F, trace="none",
#            col=colorpanel(100, "black", "white"),
#            #ColSideColors=mycols[condition], RowSideColors=mycols[condition],
#            margin=c(10, 10), main="Sample Distance Matrix")
#  legend("topleft", legend=unique(condition), col=mycols,lty=1,lwd=5,cex=0.4)
```

https://sebastianraschka.com/Articles/heatmaps_in_r.html

```{r}
my_palette <- colorRampPalette(c("blue", "green", "yellow","white"))(n = 399)
my_palette <- colorRampPalette(c("blue4", "darkcyan", "darkseagreen3","lemonchiffon1","white"))(n = 99)

# my_palette <- rev(brewer.pal(9,"YlGnBu"))
col_breaks = c(seq(0,30,length=20),  # for red
  seq(31,44,length=20),           # for yellow
  seq(45,57,length=20),
  seq(58,70,length=20),
  seq(71,100,length=20)) 
# col_breaks = c(seq(1,100, length=10) )
mat_data <- data.matrix(apply(my_data_VO2[,-11][-7,],2,function(x){round(as.numeric(x)/as.numeric(max(x))*100,2)}))
rownames(mat_data) <- rownames(my_data_VO2[-7,])
heatmap.2(mat_data[order(mat_data[,10]),],
  # cellnote = F,  # same data set for cell labels
  main = "Correlation", # heat map title
  notecol="black",      # change font color of cell labels to black
  density.info="none",  # turns off density plot inside color legend
  trace="none",         # turns off trace lines inside the heat map
  margins =c(5,5),     # widens margins around plot
  col=my_palette,       # use on color palette defined earlier
  breaks=col_breaks,    # enable color transition at specified limits
  dendrogram="none",     # only draw a row dendrogram
  Rowv="NA",
  Colv=T,
  colsep=0:ncol(mat_data),
  rowsep=0:nrow(mat_data),
  sepcolor="gray70",
  sepwidth=c(0.01,0.01),)
```

## Study different T values

* I will run the DL and DM prediction with different T values and store the correlations in a variable.
* Next, I will plot the 3 variables in a contour plot or a plot of T vs R_DL*R_DM, for instance.
```{r}
get_cor <- function(pData, xVar, yVar) {
  model_cor <- cor(pData[,yVar],pData[,xVar])**2
  return(round(model_cor,2))
}
```

```{r}
f_PM <- function(t,y,DM,params){
  dy <- - DM / (as.numeric(params[2])*ct_T*(0.0032*y[1]+1.4*as.numeric(params[1])*as.numeric(params[3]))) * (y[1]-as.numeric(params[4]))
  list(dy)
}
f_PL <- function(t,y,DL,params){
  dy <- DL / (as.numeric(params[2])*ct_T*(0.0032*y[1]+1.4*as.numeric(params[1])*as.numeric(params[3]))) * (as.numeric(params[4])-y[1])
  list(dy)
}
```


```{r}
calc_PM <- function (data) {
  hb <- data[6]
  q <- data[10]
  satv <- as.numeric(data[8])/100
  pmito <- data[15]
  mysol <- bvpshoot(yini=as.numeric(data[4]),yend=as.numeric(data[5]),x=t,func=f_PM,extra=50,maxiter=100000000,params=c(hb,q,satv,pmito))
  DM <- attr(mysol,"roots")
  plot(mysol[,1],mysol[,2],type="l")
  return(DM$root/10)
}
```

```{r}
calc_PL <- function (data) {
  hb <- data[6]
  q <- data[10]
  satv <- as.numeric(data[8])/100
  pao2 <- data[13]
  mysol <- bvpshoot(yini=as.numeric(data[5]),yend=as.numeric(data[4]),x=t,func=f_PL,extra=20,maxiter=100000000,params=c(hb,q,satv,pao2))
  DL <- attr(mysol,"roots")
  plot(mysol[,1],mysol[,2],type="l")
  return(DL$root/10)
}
```

```{r}
#ct_T=0.00833
#ct_T=0.01
dmcor <- NULL
dlcor <- NULL
dl7cor <- NULL
for (ct_T in seq(0.001,0.01,0.0018325)){
  print(ct_T)
  t <- seq(0,ct_T,ct_T/20)
  par(mfrow=c(4,6))
  my_data_VO2$DM <- round(apply(my_data_VO2,1,calc_PM),2)
  # print(my_data_VO2$DM)
  par(mfrow=c(4,6))
  my_data_VO2$DL <- round(apply(my_data_VO2,1,calc_PL),2)
  # print(my_data_VO2$DL)
  dmcor <- c(dmcor,get_cor(my_data_VO2,"VO2","DM"))
  dlcor <- c(dlcor,get_cor(my_data_VO2,"VO2","DL"))
  dl7cor <- c(dl7cor, get_cor(my_data_VO2[-7,],"VO2","DL"))
}
```
## Check results

Agus suggested me to manually check some of the results:
```
Agus 12:04
if you put the DL value from the output and calculate using the formula on 
the right hand side, do you get the outputtted derivative?
```

```{r}
#Example with first Row
#plot as in formula
calc_PL <- function (data) {
  hb <- data[6]
  q <- data[10]
  satv <- as.numeric(data[8])/100
  pao2 <- data[13]
  mysol <- bvpshoot(yini=as.numeric(data[5]),yend=as.numeric(data[4]),x=t,func=f_PL,extra=20,maxiter=100000000,params=c(hb,q,satv,pao2))
  DL <- attr(mysol,"roots")
  plot(mysol[,1],mysol[,2],type="l")
  return(mysol[,2])
}
DLfirst <- calc_PL(my_data_VO2[1,])
# x <- t
x <- DLfirst
get_DL <- function (x,data){
  DL <- data[16]
  hb <- data[6]
  q <- data[10]
  satv <- as.numeric(data[8])/100
  pao2 <- data[13]
  y <- DL / (q*ct_T*(0.0032*x+1.4*hb*satv)) * (pao2-x)
  return(y)
}
# get_DL(x[2],my_data_VO2[1,],DLfirst)
k <- my_data_VO2[1,]
s <- as.numeric(unlist(lapply(x,get_DL,data=k)))
```




```{r echo =F}
################################################
#EXAMPLES FOR BVPSOLVE
```

```{r echo =F}
f <- function(x,y,parms){
  dy <- y[2]
  dy2 <- -4*y[1]
  list(c(dy,dy2))
}
x <- seq(0,pi/4,0.01)
mysol <- bvptwp(yini=c(-2,NA),yend=c(10,NA),x=x,func=f)
k<-function(x){-2*cos(2*x)+10*sin(2*x)}
k2 <- function(x){4*sin(2*x)+20*cos(2*x)}
par(mfrow=c(2,2))
plot(mysol[,1],mysol[,2],type="l")
plot(x,k(x),type="l")
plot(mysol[,1],mysol[,3],type="l")
plot(x,k2(x),type="l")
```


sensitivity analysis for the T
plot DAVo2 vs Q (calculate DAVO2 as difference of VO2 in arteria and veins)

