---
title: "GSE304028 Microarray Analysis"
output: html_notebook
---
Processing data for Helena Qin
Study of differentially expressed genes and pathway analysis
Is ANXA1 present in any of them?

Data and code from:
https://www.ncbi.nlm.nih.gov/geo/geo2r/?acc=GSE30428
https://rawgit.com/bioinformatics-core-shared-training/microarray-analysis/master/affymetrix.nb.html

LOAD LIBRARIES
```{r}
library(oligo)
library(GEOquery)
library(limma)
library(RColorBrewer)
library(cluster)
library(biomaRt)
library(purrr)
library(gplots)
library(genefilter)
library(lattice)
library(mogene10sttranscriptcluster.db)
```

READ INPUT
ADD INFO

```{r}
targetsFile <- "phenodata_OK.txt"
pd <- read.AnnotatedDataFrame(targetsFile,header=TRUE,sep="",row.names=1)
# pData(pd)
rawData <- read.celfiles(rownames(pData(pd)))
filename <- sampleNames(rawData)
pData(rawData)$filename <- filename
sampleNames <- sub(".*_", "", filename)
sampleNames <- sub(".CEL.gz$", "", sampleNames)
sampleNames(rawData) <- sampleNames
pData(rawData)$group <- ifelse(grepl("PAC", sampleNames(rawData)),
                               "CASE", "Control")
pData(rawData)
```

QC ANALYSIS: plot raw array data

```{r}
# par(mfrow=c(2,2))
# layout(matrix(c(1,2,3,4), 2, 2, byrow = TRUE))
image(rawData[,1])
image(rawData[,2])
image(rawData[,3])
image(rawData[,4])
```
NORMALIZE reading data
```{r}
normData <- rma(rawData)
```

COMPARE RAW WITH NORM DATA

MAPLOT RAW
```{r}
MAplot(rawData[, 1:4], pairs=TRUE)
```
MAPLOT NORM
```{r}
MAplot(normData[, 1:4], pairs=TRUE)
```
BOXPLOT AND HIST RAW VS NORM
```{r}
par(mfrow=c(1,2))
boxplot(rawData,target='core',main="raw")
boxplot(normData,main="norm")
```
```{r}
hist(rawData,target="core")
hist(normData)
```

LOAD PREPROCESSED DATA FROM GEO TO REPRODUCE RESULTS
SERIES MATRIX LOADING...

```{r}
preprocessedData <- getGEO(filename="GSE30428_series_matrix.txt")
gsms <- "XXXXXXXX1100"
sml <- c()
for (i in 1:nchar(gsms)) { sml[i] <- substr(gsms,i,i) }
sel <- which(sml != "X")
preprocessedData <- preprocessedData[ , sel]
```

CHECK THAT RESULTS LOOK LIKE NORMALIZED DATA FROM CEL FILES
```{r}
boxplot(preprocessedData)
MAplot(preprocessedData[, 1:4], pairs=TRUE)
hist(preprocessedData)
```

Continue with downstream analysis (DE genes)
Check if 2 groups can be clustered in different clusters

```{r}
varFiltered <- varFilter (preprocessedData)
# dim (varFiltered)
# nrow (preprocessedData) / nrow (varFiltered)
euc.dist <- dist (t(exprs(varFiltered)))
#euc.dist
corMat <- cor(exprs(varFiltered))
#corMat
cor.dist <- as.dist(1 - corMat)
#cor.dist
clust <- hclust(cor.dist)
# clust
# names(clust)
# clust$order
# plot(clust)
clust.ward <- hclust(cor.dist,method = "ward.D")
# par(mfrow=c(1,2))
# plot(clust)
# plot(clust.ward)
# plot(clust,labels=paste("Sample", 1:4))
pd <- pData(preprocessedData)
# pd
#View(pd)
plot(clust,labels=pd$description)
```


check heatmap to idenfity any possible errors in replicates

```{r}
geneVar = rowSds(exprs(preprocessedData))
# sd(exprs(preprocessedData)[1,])
# geneVar[1]
# sd(exprs(preprocessedData)[2,])
# geneVar[2]
# length(geneVar)
highVarGenes = order (geneVar, decreasing = TRUE )[1:100]
SampleGroup <- c("case1","case2","ctrl1","ctrl2")
groupColours <- rep("red",length(SampleGroup))
groupColours[3:4] <- "blue"

# heatmap.2 (as.matrix(exprs(preprocessedData)[highVarGenes, order(SampleGroup)]),
#          labCol = SampleGroup[order(SampleGroup)])
```

```{r}
hmcol <- brewer.pal(11 , "RdBu")
heatmap.2 (as.matrix(exprs(preprocessedData)[highVarGenes, order(SampleGroup)]) , 
         labCol = SampleGroup[order(SampleGroup)],
         col=hmcol,scale='row')
```

```{r}
 supervised.clus <- pam(euc.dist,k=2)
 clusplot(supervised.clus)
#supervised.clus$clustering
```
 
PCA 

```{r}
pca.geneExpression <- prcomp(t(exprs(varFiltered)))
summary(pca.geneExpression)
# plot(pca.geneExpression)
# plot(pca.geneExpression$x[,1],pca.geneExpression$x[,2])
# plot(pca.geneExpression$x[,1],pca.geneExpression$x[,2],
#      pch=16,col=as.character(groupColours))
plot(pca.geneExpression$x[,1],pca.geneExpression$x[,2],
     pch=16,col=as.character(groupColours))
legend("bottomright",fill=groupColours,legend=SampleGroup)
text(pca.geneExpression$x[,1],pca.geneExpression$x[,2]-0.01,labels = pd$geo_accession)
# boxplot(pca.geneExpression$x[,1] ~ SampleGroup)

```

DE analysis using limma
create contrast matrix and run eBayes function

```{r}
pData(preprocessedData)$source_name_ch1[4] <- pData(preprocessedData)$source_name_ch1[3]
design <- model.matrix(~0+droplevels(pData(preprocessedData)$source_name_ch1))
#design
colnames(design) <- c("CASE","CONTROL")
 design
```
```{r}
fit1<-lmFit(preprocessedData, design)
contrasts <- makeContrasts(CASE - CONTROL, levels=design)
fit2 <- contrasts.fit(fit1, contrasts)
fit3 <- eBayes(fit2)
topTable(fit3)
```

Get results and find ENSEMBL code match for all probeids

```{r}
tab<-topTable(fit3, number = 3e6, adjust.method = "BH", sort.by = "p")
pr<-row.names(tab)
tab$symbol <- rapply(strsplit(tab$gene_assignment," // "), function(x) x[2])
x <- mogene10sttranscriptclusterENSEMBL
mapped_genes <- mappedkeys(x)
xx <- as.list(x[mapped_genes])
sy <- xx[pr]
head(sy,20)
tab$ensemblid<- as.character(sapply(sy, function(k){k[1]}))
# length(labels(sy)[!is.na(labels(sy))])
head(tab,20)
```

Match also with ENTREZID

```{r}
xEN <- mogene10sttranscriptclusterENTREZID
mapped_genesEN <- mappedkeys(xEN)
xxEN <- as.list(xEN[mapped_genesEN])
syEN <- xxEN[pr]
length(labels(syEN)[!is.na(labels(syEN))])
tab$entrezid<- as.character(sapply(syEN, function(k){k[1]}))
head(tab,20)
```

WRITE AND SAVE RESULTS IN CSV FORMAT FOR GSEA ANALYSIS
only top3000 genes (max in DAVID)

```{r}
outdataSUM <- as.data.frame(tab$symbol)
outdataSUM$pval <- tab$P.Value
outdataSUM$ensemblid <- tab$ensemblid
outdataSUM$logFC <- tab$logFC
write.table(outdataSUM[1:3000,],file="GSE30428_genes_out.csv",sep=",",quote=FALSE,row.names = FALSE,col.names=TRUE)
```

same for ranking (according to pvalue), to run GSEA offline

```{r}
outdataRNK <- as.data.frame(tab$symbol)
outdataRNK$pval <- tab$P.Value*-1
write.table(outdataRNK[1:3000,],file="GSE30428_genes_out.rnk",sep="\t",quote=FALSE,row.names = FALSE,col.names=FALSE)
```

Heatmap of top50 logfc expressed genes


```{r}
myGenesDF.logFCsort <- tab[order(abs(tab$logFC),decreasing=T),][,c(1,13:20)]
# myGenesDF.logFCsort[1:30,]
# match(selgenes,rownames(exprs(preprocessedData)))
# exprs(preprocessedData)[28158,]
selgenes <- as.numeric(myGenesDF.logFCsort$ID[1:30])
selgenes.sym <- myGenesDF.logFCsort$symbol[1:30]
bk = unique(c(seq(-2,-0.5, length=50),seq(-0.5,0.5, length=50), seq(0.5,2,length=50)))
hmcol<- colorRampPalette(c("white","red2", "red4"))(length(bk)-1)
heatmap.2 (as.matrix(exprs(preprocessedData))[match(selgenes,rownames(exprs(preprocessedData))),],labCol = SampleGroup[order(SampleGroup)],col=hmcol,labRow=selgenes.sym,trace='none',scale='row')
```

Expression levels of ANXA1 in all datasets
```{r}
mygene_expr <- "Anxa1"
gene_row <- match(mygene_expr,tab$symbol)
print(tab[gene_row,c(1,13:20)])
probeid_row <- tab$ID[gene_row]
probe_EXPR_row <- match(probeid_row,rownames(exprs(preprocessedData)))
dotplot(as.matrix(exprs(preprocessedData)[probe_EXPR_row,]),ylab=SampleGroup,xlab=paste("MicroArray Raw expression level - ", mygene_expr),col=groupColours,groups=SampleGroup)
```

VOLCANO PLOT OF THE SAME DATASET HIGHLIGHTING THE TOP LOGFC

```{r}
volcanoplot(fit3,highlight=40,names=rapply(strsplit(fit3$genes$gene_assignment," // "), function(x) x[2]))
```

PATHWAY ANALYSIS

```{r}
# library(gage)
# library(pathview)
# 
# data(kegg.gs)
# exp.fc=tab$logFC
# names(exp.fc)=tab$entrezid
# fc.kegg.p <- gage(exp.fc, gsets = kegg.gs, ref = NULL, samp = NULL)
# sel <- fc.kegg.p$greater[, "q.val"] < 0.1 &         !is.na(fc.kegg.p$greater[, "q.val"])
# path.ids <- rownames(fc.kegg.p$greater)[sel]
# sel.l <- fc.kegg.p$less[, "q.val"] < 0.1 &            !is.na(fc.kegg.p$less[,"q.val"])
# path.ids.l <- rownames(fc.kegg.p$less)[sel.l]
# path.ids2 <- substr(c(path.ids, path.ids.l), 1, 8)
# head(path.ids)
```

Plot of pathways

```{r}
# pv.out.list <- sapply(path.ids2[1:3], function(pid) pathview(gene.data =  exp.fc, pathway.id = pid,                      species = "hsa", out.suffix="limma"))
# library(png)
# 
# files <- list.files(pattern = 'limma.png')
# for (i in files)
# {
# img <- readPNG(i)
# grid::grid.raster(img)
# }
# 
# # img <- readPNG(files[1])
# # grid::grid.raster(img)
```

