---
title: "Training Peaks"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    theme: cosmo
    df_print: paged
    highlight: tango
    # code_folding: hide
    # fig_width: 12
    # fig_height: 12
# output:
#   epuRate::BAKER:
#     toc: TRUE
#     number_sections: FALSE
#     code_folding: "show"
---

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

***

# TrainingPeaks Data

## Preliminary tasks
### Load Libraries

Load libraries for reading fit data and create maps and plots.

```{r warning=FALSE, results='hide', echo='FALSE'}
library(fit)
library(ggplot2)
library(leaflet)
library(dplyr)
# library(cetcolor)
library(stringr)
```

*** 

# Activity per athlete

> **Note**
>
> We are missing personal data from all patients (Name, VO2max, Weight, etc.)

## Summarize activities per athlete

Number of activities per category?

* Month 2,12 year 0,2
* Full activities for the 2 years?
* N activities cycling, running, other?
* Heart data?


### Sport type definition

  * https://developer.garmin.com/downloads/connect-iq/monkey-c/doc/Toybox/ActivityRecording.html

Specify main sports with data (running, cycling, swimming, rowing)
```{r}
sport_code <- c(0,1,2,5,10,15)
names(sport_code) <- c("Undefined","Running","Cycling","Swimming","Cardio","Rowing")
```
  
### Selected Athletes

Correct naming

```{r}
sel.athletes <- c("LEUVEN_PH/LOICSEGAERT","HASSELT_PH/RUBENAPERS","HASSELT_PH/SANDERDEPESTEL","HASSELT_PH/ARNEMARIT",
                  "HASSELT_PH/AARONVANPOUCKE","AUS_PH/FERGUSSULLY","AUS_PH/MATHEWROSS","AUS_PH/CYRUSMONK","AUS_PH/BENTLEYOLDEN",
                  "AUS_PH/DAVIDRANDALL","HASSELT_PH/WARDVANHOOF","LEUVEN_PH/ROBSCHEERLINCK","LEUVEN_PH/TRISTANGERRYN")
```


```{r}
### FOR TESTING
# sel.dirs[50]
# dataheart <- try(read.fit(paste0(sel.dirs[50],"/",list.files(sel.dirs[50])[1])),silent=T)
# get_activity_info_from_fitdata(dataheart,training.all = training.all, ath.id = "BA_007")
```


## Define Functions

Create a function to calculate all predictors as previously, given a list of Fit files:

```{r}
get_activity_info_from_fitdata <- function (fitdata,training.all=training.all,ath.id) {
  #initialize activity summary
  activity <- as.data.frame(matrix(NA,ncol=1))
  colnames(activity) <- "date"
  activity.date <- get_date_GARMIN(fitdata$session$start_time)
  activity$date <- format(activity.date,"%Y%m%d")
  activity$year <- format(activity.date,"%Y")
  activity$month <- format(activity.date,"%m")
  activity$week <- format(activity.date,"%W")
  activity$start_time <- format(activity.date,"%H%M")
  activity$time.min <- round((max(fitdata$record$timestamp)-fitdata$record$timestamp[1])/60,2)

  if (any(names(fitdata$record) %in% "heart_rate")){
    activity$hrmax.activity <- max(fitdata$record$heart_rate[1:(length(fitdata$record$heart_rate)-3)],na.rm=T)
    #check if there is heart_rate info (not in swimming, surfing and other watersports)
    if(is.null(fitdata$zones_target$max_heart_rate)){
      hrmax.prev <- if(is.null(training.all$ath_id)){0}else{training.all %>% filter(ath_id==ath.id) %>% summarize(max(hrmax.athlete)) %>% as.numeric()}
      activity$hrmax.athlete <- max(min(round(activity$hrmax.activity*1.03,0),199),hrmax.prev,na.rm=T)
    } else {
      activity$hrmax.athlete <- fitdata$zones_target$max_heart_rate
    }
    activity$hrmax.perc <- round(activity$hrmax.activity/activity$hrmax.athlete*100,2)
    activity$hrmax.intensity <- activity$hrmax.perc * activity$time.min
    hr.zones <- quantile(c(0:activity$hrmax.athlete),probs=seq(0,1,by=0.1))
    fitdata$record$hr.zones <- findInterval(fitdata$record$heart_rate,hr.zones[6:10])
    hr.zones.table<-round(table(fitdata$record$hr.zones)/sum(table(fitdata$record$hr.zones))*100,2)
    hr.zones.table[c("0","1","2","3","4","5")[!c("0","1","2","3","4","5") %in% names(hr.zones.table)]] <- 0
    activity$hr.z1 <- hr.zones.table['1'] + hr.zones.table['0']
    activity$hr.z2 <- hr.zones.table['2']
    activity$hr.z3 <- hr.zones.table['3']
    activity$hr.z4 <- hr.zones.table['4']
    activity$hr.z5 <- hr.zones.table['5']
    activity$hr.avg <- fitdata$session$avg_heart_rate
  } else {
    activity$hrmax.activity <- NA
    #check if there is heart_rate info (not in swimming, surfing and other watersports)
    if(is.null(fitdata$zones_target$max_heart_rate)){
      hrmax.prev <- if(is.null(training.all$ath_id)){0}else{training.all %>% filter(ath_id==ath.id) %>% summarize(max(hrmax.athlete)) %>% as.numeric()}
      activity$hrmax.athlete <- if(hrmax.prev == 0){NA}else{hrmax.prev}
    } else {
      activity$hrmax.athlete <- fitdata$zones_target$max_heart_rate
    }
    # activity$hrmax.athlete <- NA
    activity$hrmax.perc <- NA
    activity$hrmax.intensity <- NA
    activity$hr.z1 <- NA
    activity$hr.z2 <- NA
    activity$hr.z3 <- NA
    activity$hr.z4 <- NA
    activity$hr.z5 <- NA
    activity$hr.avg <- NA
  }
  # VT zones
  if (any(names(fitdata$record) %in% "heart_rate")){
    vt.zones <- quantile(c(0:activity$hrmax.athlete),probs=c(0,0.83,0.94,1))
    fitdata$record$vt.zones <- findInterval(fitdata$record$heart_rate,vt.zones)
    vt.zones.table <- round(table(fitdata$record$vt.zones)/sum(table(fitdata$record$vt.zones))*100,2)
    vt.zones.table[c("1","2","3")[!c("1","2","3") %in% names(vt.zones.table)]] <- 0
    activity$vt.z1 <- vt.zones.table['1'] 
    activity$vt.z2 <- vt.zones.table['2']
    activity$vt.z3 <- vt.zones.table['3']
  } else {
    # activity$hrmax.athlete <- NA
    activity$vt.z1 <- NA
    activity$vt.z2 <- NA
    activity$vt.z3 <- NA
  }
  activity$sport_code <- fitdata$session$sport
  activity$sport_type <- names(sport_code)[sport_code == activity$sport_code]
  activity$speed.avg <- round(fitdata$session$avg_speed*3.79,1)
  activity$cal <- if(is.null(fitdata$session$total_calories)){NA}else{fitdata$session$total_calories}
  activity$ascent <- if(is.null(fitdata$session$total_ascent)){NA}else{fitdata$session$total_ascent}
  activity$power.max <- if(is.null(fitdata$session$max_power)){NA}else{fitdata$session$max_power}
  activity$power.avg <- if(is.null(fitdata$session$avg_power)){NA}else{fitdata$session$avg_power}
  activity$power.norm <- if(is.null(fitdata$session$normalized_power)){NA}else{fitdata$session$normalized_power}
  activity$work <- if(is.null(fitdata$session$total_work)){NA}else{fitdata$session$total_work}
  activity$stress.score <- if(is.null(fitdata$session$training_stress_score)){NA}else{fitdata$session$training_stress_score}
  activity$total.dist <- fitdata$session$total_distance
  activity$intensity.factor <-  if(is.null(fitdata$session$intensity_factor)){NA}else{fitdata$session$intensity_factor}
  activity$training.effect <-  if(is.null(fitdata$session$total_training_effect)){NA}else{fitdata$session$total_training_effect}
  # print(all.activity)
  return(activity)
}
```


Get date
https://stat.ethz.ch/R-manual/R-devel/library/base/html/strptime.html  
http://www.nlvocables.com/blog/?p=969 (garmin timestamp)

```{r}
get_date_GARMIN <- function(tstmp){
  as.POSIXct(tstmp, origin="1989-12-31",tz="UTC")
}
# data.date <- get_date_GARMIN(data$session$start_time)
# data.date.day <- format(data.date,"%Y%m%d")
# data.date.start_time <- format(data.date,"%H%M")
# data.date.year <- format(data.date,"%Y")
# data.date.month <- format(data.date,"%m")
# data.date.week <- format(data.date,"%W")
```

## Get directories


```{r}
dirs.ph <- list.dirs("PRO_HEART")
sel.dirs <- dirs.ph[str_detect(dirs.ph,paste(sel.athletes,collapse="|"))]
sel.dirs <- sel.dirs[str_detect(sel.dirs,"YEAR")]
# sel.dirs
```

Initialize athlete list

```{r}
ath.info<- as.data.frame(unique(do.call(rbind, str_split(sel.dirs,"/"))[,3]),stringsAsFactors =FALSE)
colnames(ath.info) <- "name"
ath.info$ath_id <- paste0("BA_",str_pad(as.numeric(as.factor(ath.info$name)),3,pad="0"))
ath.info$total_activities <- 0
ath.info$month_2_year_0 <- 0
ath.info$month_12_year_0 <- 0
ath.info$month_2_year_2 <- 0
ath.info$month_12_year_2 <- 0
ath.info$full_2years <- 0
ath.info$n_cycling <- 0
ath.info$n_running <- 0
ath.info$n_other <- 0
ath.info$n_heart <- 0
ath.info$perc_heart <- 0
```

# Activity sessions

```{r}
error.sessions <- NULL
dup.sessions <- setNames(data.frame(matrix(ncol = 3, nrow = 0)), c("dup_file", "original_file","reason"))
training.all <- NULL
for (dir in sel.dirs){
  print(dir)
  dir.name <- unlist(str_split(dir,"/"))[3]
  dir.monthcategory <- tolower(unlist(str_split(dir,"/"))[4])
  dir.files <- list.files(dir,pattern="fit")
  for (file in dir.files) {
    file.fullpath <- paste0(dir,"/",file)
    # check if file has already been processed in a different folder (some cases with full/monthly divisions)
    if (any(sapply(str_split(training.all$file,"/"),function(x) x[5]) == file & training.all$ath_id == ath.info$ath_id[ath.info$name==dir.name])){
      dup.source <- as.character(training.all$file[sapply(str_split(training.all$file,"/"),function(x) x[5]) == file & training.all$ath_id == ath.info$ath_id[ath.info$name==dir.name]])
      dup.sessions <- rbind(dup.sessions,setNames(data.frame(matrix(c(file.fullpath,dup.source,"rep_file"),nrow=1)),c("dup_file", "original_file","reason")))
      next
    }
    # read fit file
    fitdata <- try(read.fit(paste0(dir,"/",file)),silent=T)
    # skip files with errors
    if (class(fitdata) == "try-error"){
      error.sessions <- c(error.sessions,paste0(dir,"/",file))
      next
    }
    # get session data
    session <- try(get_activity_info_from_fitdata(fitdata,training.all = training.all, ath.id = ath.info$ath_id[ath.info$name==dir.name]), silent=T)
    # skip files with errors
    if (class(session) == "try-error"){
      error.sessions <- c(error.sessions,paste0(dir,"/",file))
      next
    }
    # check if activity is already present on the db and skip if so
    if (any(training.all$date == session$date & training.all$start_time == session$start_time & training.all$ath_id == ath.info$ath_id[ath.info$name==dir.name] & training.all$time.min == session$time.min)){
      dup.source <- as.character(training.all$file[training.all$date == session$date & training.all$start_time == session$start_time & training.all$ath_id == ath.info$ath_id[ath.info$name==dir.name] & training.all$time.min == session$time.min])
      dup.sessions <- rbind(dup.sessions,setNames(data.frame(matrix(c(file.fullpath,dup.source,"rep_activity"),nrow=1)),c("dup_file", "original_file","reason")))
      next
    }
    # update session to be counted and add the info to training.all df
    # ath.info[ath.info$name == dir.name,dir.monthcategory] <-  ath.info[ath.info$name == dir.name,dir.monthcategory] + 1
    training.all <- rbind(training.all,cbind(ath_id=ath.info$ath_id[ath.info$name == dir.name],session,file=file.fullpath))
    # # check sport type
    fitdata.sporttype <- fitdata$session$sport
    if (fitdata.sporttype == 2) {
      ath.info[ath.info$name == dir.name,"n_cycling"] <- ath.info[ath.info$name == dir.name,"n_cycling"] + 1
    } else if (fitdata.sporttype == 1) {
      ath.info[ath.info$name == dir.name,"n_running"] <- ath.info[ath.info$name == dir.name,"n_running"] + 1
    } else {
      ath.info[ath.info$name == dir.name,"n_other"] <- ath.info[ath.info$name == dir.name,"n_other"] + 1
    }
    # update total after each directory
    ath.info[ath.info$name == dir.name,dir.monthcategory] <- ath.info[ath.info$name == dir.name,dir.monthcategory] + 1
    ath.info[ath.info$name == dir.name,"total_activities"] <- ath.info[ath.info$name == dir.name,]$month_2_year_0 +
      ath.info[ath.info$name == dir.name,]$month_12_year_0 + ath.info[ath.info$name == dir.name,]$month_2_year_2 +
      ath.info[ath.info$name == dir.name,]$month_12_year_2 + ath.info[ath.info$name == dir.name,]$full_2years
    # check heart rate if present
    if (!is.null(fitdata$record$heart_rate)) {
      ath.info[ath.info$name == dir.name,"n_heart"] <- ath.info[ath.info$name == dir.name,"n_heart"] + 1
      ath.info[ath.info$name == dir.name,"perc_heart"] <- round(ath.info[ath.info$name == dir.name,"n_heart"] /
                                                                  ath.info[ath.info$name == dir.name,"total_activities"] * 100,2)
    }
  }
}
```

```{r}
ath.info
```

```{r}
save(ath.info,training.all,error.sessions,dup.sessions,file=
       paste0("trainingpeaks_data_",format(as.Date(Sys.Date()),format="%y%m%d"),".rda"))
```



* Number of Activities
* Volume (Time)
* Intensity
  * %HRmax
  * Thresholds (Time in T1/T2)
  * Volume x %HRmax
* Sessions per week
* Average time
* Average speed (cycling vs running, etc?)
  * Event type = 1 == cycling??
  * sport = 2?
    * 2 = cycling
    * 5 = swimming
  * https://developer.garmin.com/downloads/connect-iq/monkey-c/doc/Toybox/ActivityRecording.html
* Average HR
* Heart beats per week/year of training
* Average accumulated power
* Data from TrainingPeaks? (data$session)
  * Calories
  * Total ascent
  * max_power
  * average_power / normalized_power
  * total work
  * training_stress_score
  * total_distance
  * intensity_factor
  * avg_HR

  * Average accumulated power (ATHLETE CALCULATION)
* Heart beats per week/year of training (ATHLETE CALCULATION)
* Data from TrainingPeaks? (data$session)
  * Calories

  * Thresholds (Time in Z5/Z4/Z3)
https://blogs.sas.com/content/efs/2018/01/26/data-driven-fitness-vo2-max-lactate-threshold-heart-rate/


Summarize power and add to ath.info

```{r}
load('trainingpeaks_data_200630.rda')

cohort_vs_athlete <- as.data.frame(t(do.call("cbind",str_split(str_remove(str_remove(sel.dirs[str_detect(sel.dirs,"FULL")],"PRO_HEART/"),"/FULL_2YEARS"),"/"))))
ath.info.ok <- merge(merge(ath.info, cohort_vs_athlete,by.x="name",by.y="V2"),
                     training.all %>%
                       group_by(ath_id) %>%
                       summarize(n_power=sum(!is.na(power.avg)),n_heart_power=sum(!is.na(power.avg) & !is.na(hrmax.activity))),by='ath_id') %>%
  mutate(cohort = V1,perc_power = round(n_power/total_activities*100,2), perc_heart_power = round(n_heart_power/total_activities*100,2)) %>%
  select(-V1)


# training.all %>%
#   filter(ath_id=="BA_003",!is.na(hrmax.activity))
write.csv(ath.info.ok, file=paste0("tpeaks_summary_athletes_",format(as.Date(Sys.Date()),format="%y%m%d"),".csv"))
write.csv(training.all, file=paste0("tpeaks_summary_activities_",format(as.Date(Sys.Date()),format="%y%m%d"),".csv"))
write.csv(error.sessions, file=paste0("tpeaks_summary_error_",format(as.Date(Sys.Date()),format="%y%m%d"),".csv"))
write.csv(dup.sessions, file=paste0("tpeaks_summary_dups_",format(as.Date(Sys.Date()),format="%y%m%d"),".csv"))
```

```{r}
# for (i in ath.info$name) { print(i);print(sum(!is.na(str_extract(error.sessions,i))))}
```

```{r}
for (i in ath.info$name) {
  # print(i);
  print(sum(!is.na(str_extract(dup.sessions$dup_file,i))))
}
```


```{r}
# nfiles <- as.data.frame(matrix(rep(0,1)))
# for (dir in sel.dirs){
#   dir.name <- unlist(str_split(dir,"/"))[3]
#   nfiles[`dir.name`] <- 0
# }
# nfiles <- nfiles %>% select(-V1)
# 
# for (dir in sel.dirs){
#   # print(dir)
#   dir.name <- unlist(str_split(dir,"/"))[3]
#   dir.monthcategory <- tolower(unlist(str_split(dir,"/"))[4])
#   dir.files <- list.files(dir)
#   nfiles[`dir.name`] <- nfiles[`dir.name`]+length(dir.files)
# }
```

duplicate check 

```{r}
# dupfile <- 'PRO_HEART/AUS_PH/BENTLEYOLDEN/MONTH_12_YEAR_2/bentleyno.2018-12-10-07-07-12-257Z.GarminPush.27656281043.fit'
# # dupfile <- 'PRO_HEART/AUS_PH/BENTLEYOLDEN/FULL_2YEARS/bentleyno.2018-06-21-11-57-17-252Z.GarminPush.21832615345.fit'
# dir.name <-   unlist(str_split(dupfile,"/"))[3]
# fitdata <- read.fit(dupfile)
# session <- get_activity_info_from_fitdata(fitdata,training.all = training.all, ath.id = ath.info$ath_id[ath.info$name==dir.name])
# session$file <- dupfile
# rbind(session,training.all[which(training.all$date == session$date & training.all$start_time == session$start_time & training.all$ath_id == ath.info$ath_id[ath.info$name==dir.name] & training.all$time.min == session$time.min),] %>% select(-ath_id))
```

year summary

```{r}
for (i in ath.info$name) {
  print(i);
  k <- training.all %>% group_by(ath_id,year) %>% tally() %>% filter(ath_id==ath.info[ath.info$name == i,"ath_id"]) %>% mutate(n=as.numeric(n)) %>% t() %>% data.frame()
  print(k)
}
```

***
<!-- <!-- -->

<!-- ## Superpredictors as summary -->

<!-- Superpredictors are averages or totals of each of the activity predictors, as a way to summarize training data for each athlete during the selected period. -->
<!-- Here we calculate all of them to compare with vo2max in the final step. -->

<!-- ### Definition -->

<!-- Define function to: -->
<!-- * Initialize variables -->

<!-- ```{r} -->
<!-- # final.data <- vo2max -->
<!-- prep_superpredictors <- function(final.data){ -->
<!--   final.data$activities.total <- NA -->
<!--   final.data$time.total <- NA -->
<!--   final.data$time.avg <- NA -->
<!--   final.data$hrmax.perc.avg <- NA -->
<!--   final.data$hrmax.max <- NA -->
<!--   final.data$hrmax.intensity.total <- NA -->
<!--   final.data$hrmax.intensity.avg <- NA -->
<!--   final.data$hr.z5.avg <- NA -->
<!--   final.data$hr.z4.avg <- NA -->
<!--   final.data$hr.z3.avg <- NA -->
<!--   final.data$hr.z2.avg <- NA -->
<!--   final.data$hr.z1.avg <- NA -->
<!--   final.data$hr.avg <- NA -->
<!--   final.data$cal.total <- NA -->
<!--   final.data$cal.avg <- NA -->
<!--   final.data$power.avg <- NA -->
<!--   final.data$work.avg <- NA -->
<!--   final.data$stress.score.avg <- NA -->
<!--   final.data$dist.total <- NA -->
<!--   final.data$dist.avg <- NA -->
<!--   final.data$intensity.factor.avg <- NA -->
<!--   return(final.data) -->
<!-- } -->
<!-- ``` -->

<!-- * Calculate variables -->

<!-- ```{r} -->
<!-- # ath <- 'Liam' -->
<!-- get_superpredictors <- function (ath.data,ath.predictors,ath.name) { -->
<!--   ath.data[ath.data$name == ath.name,]$activities.total <- nrow(ath.predictors) -->
<!--   ath.data[ath.data$name == ath.name,]$time.total <- sum(ath.predictors$time.min) -->
<!--   ath.data[ath.data$name == ath.name,]$time.avg <- mean(ath.predictors$time.min,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$hrmax.perc.avg <- mean(ath.predictors$hrmax.perc,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$hrmax.max <- if (sum(!is.na(ath.predictors$hrmax.activity)) == 0 ){NA}else{max(ath.predictors$hrmax.activity,na.rm=T)} -->
<!--   ath.data[ath.data$name == ath.name,]$hrmax.intensity.total <- sum(ath.predictors$hrmax.intensity,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$hrmax.intensity.avg <- mean(ath.predictors$hrmax.intensity,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$hr.z5.avg <- mean(ath.predictors$hr.z5,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$hr.z4.avg <- mean(ath.predictors$hr.z4,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$hr.z3.avg <- mean(ath.predictors$hr.z3,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$hr.z2.avg <- mean(ath.predictors$hr.z2,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$hr.z1.avg <- mean(ath.predictors$hr.z1,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$hr.avg <- mean(ath.predictors$hr.avg,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$cal.total <- sum(ath.predictors$cal,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$cal.avg <- mean(ath.predictors$cal,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$power.avg <- mean(ath.predictors$power.avg,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$work.avg <- mean(ath.predictors$work,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$stress.score.avg <- mean(ath.predictors$stress.score,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$dist.total <- sum(ath.predictors$total.dist,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$dist.avg <- mean(ath.predictors$total.dist,na.rm=T) -->
<!--   ath.data[ath.data$name == ath.name,]$intensity.factor.avg <- mean(ath.predictors$intensity.factor,na.rm=T) -->
<!--   ath.data[is.na(ath.data)] <- NA -->
<!--   return(ath.data) -->
<!-- } -->
<!-- ``` -->


<!-- ### Calculate Superpredictors -->

<!-- Combine the reading of the files with the superpredictor calculation -->

<!-- ```{r} -->
<!-- dirs <- list.dirs("SELECTED_FILES",recursive=F) -->
<!-- ath.superpreds <- prep_superpredictors(vo2max) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- for (ath in vo2max$name){ -->
<!-- # for (ath in 'Liam'){ -->
<!--   ath.folder <- paste0(dirs[grep(tolower(ath),tolower(dirs))],"/09/") -->
<!--   myfiles <- paste0(ath.folder, list.files(paste0(ath.folder))) -->
<!--   #get predictors -->
<!--   ath.predictors <- calc_predictors(myfiles) -->
<!--   #get superpredictors -->
<!--   ath.superpreds <- get_superpredictors(ath.superpreds,ath.predictors,ath) -->
<!-- } -->
<!-- ``` -->


<!-- And the final results for all the athletes: -->

<!-- ```{r} -->
<!-- ath.superpreds -->
<!-- ``` -->







<!-- *** -->

<!-- ## Make correlation with VO2max -->

<!-- Correlate all superpredictors with VO2max. -->
<!-- Each point corresponds to an athlete: -->

<!-- ```{r} -->
<!-- # models <- as.data.frame(matrix(NA,ncol=22)) -->
<!-- # models <- rbind(models,NA) -->
<!-- # row.names(models) <- c("rsquare","pvalue") -->
<!-- # colnames(models) <- names(ath.superpreds)[c(5,7:dim(ath.superpreds)[2])] -->
<!-- ``` -->


<!-- ```{r} -->
<!-- # for (mod in names(models)){ -->
<!-- #   md <- paste("VO2max ~ ",mod,sep = "") -->
<!-- #   lmTemp = lm(md, data = ath.superpreds) #Create the linear regression -->
<!-- #   plot(ath.superpreds[[mod]],ath.superpreds$VO2max, pch = 16, col = "blue",xlab=mod,ylab="VO2max") #Plot the results -->
<!-- #   abline(lmTemp) #Add a regression line -->
<!-- #   # print(summary(lmTemp)) -->
<!-- #   models[[mod]] <- c(summary(lmTemp)$adj.r.squared,summary(lmTemp)$coefficients[2,4]) -->
<!-- # } -->
<!-- ``` -->

<!-- * Model with all variables -->

<!-- ```{r} -->
<!-- # md <- paste("VO2max ~ ",paste(names(models)[1:3],collapse="+"),sep = "") -->
<!-- # lmTemp = lm(md, data = ath.superpreds) #Create the linear regression -->
<!-- # plot(ath.superpreds[[mod]],ath.superpreds$VO2max, pch = 16, col = "blue",xlab=mod,ylab="VO2max") #Plot the results -->
<!-- # abline(lmTemp) #Add a regression line -->
<!-- # # print(summary(lmTemp)) -->
<!-- # models$ALL <- c(summary(lmTemp)$adj.r.squared,summary(lmTemp)$coefficients[2,4]) -->
<!-- # summary(lmTemp) -->
<!-- # models -->
<!-- ``` -->




<!-- ```{r} -->
<!-- # library(purrr) -->
<!-- # vo2max <- read.csv("~/sgcluster/WORK/BIOINFO/TRAINING_PEAKS/VO2maxdata.csv") -->
<!-- # vo2max <- vo2max[vo2max$EventName == "Baseline",] -->
<!-- # row.names(vo2max) <- 1:nrow(vo2max) -->
<!-- # vo2max <- vo2max %>% mutate(name= map(strsplit(as.character(Athlete), "(?!^)(?=[[:upper:]])", perl=T),~.x[1]) %>% unlist()) -->
<!-- # vo2max -->
<!-- ``` -->

<!-- --> -->


