---
title: "Training Peaks"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    theme: cosmo
    df_print: paged
    highlight: tango
    # code_folding: hide
    # fig_width: 12
    # fig_height: 12
# output:
#   epuRate::BAKER:
#     toc: TRUE
#     number_sections: FALSE
#     code_folding: "show"
---

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

***

# TrainingPeaks Data

## Preliminary tasks
### Load Libraries

Load libraries for reading fit data and create maps and plots.

```{r warning=FALSE, results='hide', echo='FALSE'}
library(fit)
library(ggplot2)
library(leaflet)
library(dplyr)
library(cetcolor)
library(stringr)
```

*** 

# TrainingPeaks Preliminar Summary

## Summarize activities per athlete

Number of activities per category?

* Month 2,12 year 0,2
* Full activities for the 2 years?
* N activities cycling, running, other?
* Heart data?


### Sport type definition

  * https://developer.garmin.com/downloads/connect-iq/monkey-c/doc/Toybox/ActivityRecording.html

Specify main sports with data (running, cycling, swimming, rowing)
```{r}
sport_code <- c(0,1,2,5,10,15)
sport_type <- c("Undefined","Running","Cycling","Swimming","Rowing")
# predictors$sport_code <- data$session$sport
# predictors$sport_type <- sport_type[which (sport_code == predictors$sport_code)]
# predictors$speed.avg <- round(data$session$avg_speed*3.79,1)
```
  
### Athlete original data

***
**NOTE**

We are missing personal data from all patients (Name, VO2max, Weight, etc.)

***

# PRO @ HEART

## PRO@HEART in ANTWERPEN

### Athlete input files


```{r}
# dirs <- list.dirs("~/sgcluster/WORK/BIOINFO/TRAINING_PEAKS/SELECTED_FILES",recursive=F)
dirs_raw.ant <- list.dirs("PRO_HEART/ANTWERPEN_PH")
dirs.ant <- dirs_raw.ant[str_detect(dirs_raw.ant,"YEAR")]

# for (dir in dirs){
#   # k <- paste0(dirs[grep(tolower(ath),tolower(dirs))],"/09")
#   
#   print(dir)
#   print(head(list.files(dir),1))
# }
```

Initialize athlete list

```{r}
ath.ant<- as.data.frame(unique(do.call(rbind, str_split(dirs.ant,"/"))[,3]),stringsAsFactors =FALSE)
colnames(ath.ant) <- "name"
ath.ant$total_activities <- 0
ath.ant$month_2_year_0 <- 0
ath.ant$month_12_year_0 <- 0
ath.ant$n_cycling <- 0
ath.ant$n_running <- 0
ath.ant$n_other <- 0
ath.ant$n_heart <- 0
ath.ant$perc_heart <- 0
```

### Fill list

```{r}
for (dir in dirs){
  dir.name <- unlist(str_split(dir,"/"))[3]
  dir.month <- tolower(unlist(str_split(dir,"/"))[4])
  dir.files <- list.files(dir,pattern="fit")
  ath.ant[ath.ant$name == dir.name,dir.month] <-  length(dir.files)
  ath.ant[ath.ant$name == dir.name,"total_activities"] <- ath.ant[ath.ant$name == dir.name,]$month_2_year_0 + ath.ant[ath.ant$name == dir.name,]$month_12_year_0
  for (file in dir.files) {
    # read fit file
    fitdata <- read.fit(paste0(dir,"/",file))
    # check sport type
    fitdata.sporttype <- fitdata$session$sport
    if (fitdata.sporttype == 2) {
      ath.ant[ath.ant$name == dir.name,"n_cycling"] <- ath.ant[ath.ant$name == dir.name,"n_cycling"] + 1
    } else if (fitdata.sporttype == 1) {
      ath.ant[ath.ant$name == dir.name,"n_running"] <- ath.ant[ath.ant$name == dir.name,"n_running"] + 1
    } else {
      ath.ant[ath.ant$name == dir.name,"n_other"] <- ath.ant[ath.ant$name == dir.name,"n_other"] + 1
    }
    # check heart rate if present
    if (!is.null(fitdata$record$heart_rate)) {
      ath.ant[ath.ant$name == dir.name,"n_heart"] <- ath.ant[ath.ant$name == dir.name,"n_heart"] + 1
      ath.ant[ath.ant$name == dir.name,"perc_heart"] <- round(ath.ant[ath.ant$name == dir.name,"n_heart"] / ath.ant[ath.ant$name == dir.name,"total_activities"] * 100,1)
    }
  }
}
```

```{r}
ath.ant
```


*** 
## PRO@HEART in HASSELT

### Athlete input files


```{r}
# dirs <- list.dirs("~/sgcluster/WORK/BIOINFO/TRAINING_PEAKS/SELECTED_FILES",recursive=F)
dirs_raw.hasselt <- list.dirs("PRO_HEART/HASSELT_PH")
dirs.hasselt <- dirs_raw.hasselt[str_detect(dirs_raw.hasselt,"YEAR")]

# for (dir in dirs){
#   print(dir)
#   print(head(list.files(dir),1))
# }
```

Initialize athlete list

```{r}
ath.hasselt<- as.data.frame(unique(do.call(rbind, str_split(dirs.hasselt,"/"))[,3]),stringsAsFactors =FALSE)
colnames(ath.hasselt) <- "name"
ath.hasselt$total_activities <- 0
ath.hasselt$month_2_year_0 <- 0
ath.hasselt$month_12_year_0 <- 0
ath.hasselt$month_2_year_2 <- 0
ath.hasselt$month_12_year_2 <- 0
ath.hasselt$full_2years <- 0
# ath.hasselt$n_cycling <- 0
# ath.hasselt$n_running <- 0
# ath.hasselt$n_other <- 0
# ath.hasselt$n_heart <- 0
# ath.hasselt$perc_heart <- 0
```

### Fill list

```{r}
for (dir in dirs.hasselt){
  dir.name <- unlist(str_split(dir,"/"))[3]
  dir.month <- tolower(unlist(str_split(dir,"/"))[4])
  dir.files <- list.files(dir,pattern="fit")
  ath.hasselt[ath.hasselt$name == dir.name,dir.month] <-  length(dir.files)
  ath.hasselt[ath.hasselt$name == dir.name,"total_activities"] <- ath.hasselt[ath.hasselt$name == dir.name,]$month_2_year_0 + ath.hasselt[ath.hasselt$name == dir.name,]$month_12_year_0 + ath.hasselt[ath.hasselt$name == dir.name,]$month_2_year_2 + ath.hasselt[ath.hasselt$name == dir.name,]$month_12_year_2 + ath.hasselt[ath.hasselt$name == dir.name,]$full_2years
  # for (file in dir.files[1:2]) {
  #   # read fit file
  #   fitdata <- read.fit(paste0(dir,"/",file))
  #   # check sport type
  #   fitdata.sporttype <- fitdata$session$sport
  #   if (fitdata.sporttype == 2) {
  #     ath.hasselt[ath.hasselt$name == dir.name,"n_cycling"] <- ath.hasselt[ath.hasselt$name == dir.name,"n_cycling"] + 1
  #   } else if (fitdata.sporttype == 1) {
  #     ath.hasselt[ath.hasselt$name == dir.name,"n_running"] <- ath.hasselt[ath.hasselt$name == dir.name,"n_running"] + 1
  #   } else {
  #     ath.hasselt[ath.hasselt$name == dir.name,"n_other"] <- ath.hasselt[ath.hasselt$name == dir.name,"n_other"] + 1
  #   }
  #   # check heart rate if present
  #   if (!is.null(fitdata$record$heart_rate)) {
  #     ath.hasselt[ath.hasselt$name == dir.name,"n_heart"] <- ath.hasselt[ath.hasselt$name == dir.name,"n_heart"] + 1
  #     ath.hasselt[ath.hasselt$name == dir.name,"perc_heart"] <- round(ath.hasselt[ath.hasselt$name == dir.name,"n_heart"] / ath.hasselt[ath.hasselt$name == dir.name,"total_activities"] * 100,1)
  #   }
  # }
}
```

```{r}
ath.hasselt
```


*** 

## PRO@HEART in AUSTRALIA

### Athlete input files


```{r}
# dirs <- list.dirs("~/sgcluster/WORK/BIOINFO/TRAINING_PEAKS/SELECTED_FILES",recursive=F)
dirs_raw.aus <- list.dirs("PRO_HEART/AUS_PH")
dirs.aus <- dirs_raw.aus[str_detect(dirs_raw.aus,"YEAR")]

# for (dir in dirs){
#   print(dir)
#   print(head(list.files(dir),1))
# }
```

Initialize athlete list

```{r}
ath.aus<- as.data.frame(unique(do.call(rbind, str_split(dirs.aus,"/"))[,3]),stringsAsFactors =FALSE)
colnames(ath.aus) <- "name"
ath.aus$total_activities <- 0
ath.aus$month_2_year_0 <- 0
ath.aus$month_12_year_0 <- 0
ath.aus$month_2_year_2 <- 0
ath.aus$month_12_year_2 <- 0
ath.aus$full_2years <- 0
# ath.aus$n_cycling <- 0
# ath.aus$n_running <- 0
# ath.aus$n_other <- 0
# ath.aus$n_heart <- 0
# ath.aus$perc_heart <- 0
```

### Fill list

```{r}
for (dir in dirs.aus){
  dir.name <- unlist(str_split(dir,"/"))[3]
  dir.month <- tolower(unlist(str_split(dir,"/"))[4])
  dir.files <- list.files(dir,pattern="fit")
  ath.aus[ath.aus$name == dir.name,dir.month] <-  length(dir.files)
  ath.aus[ath.aus$name == dir.name,"total_activities"] <- ath.aus[ath.aus$name == dir.name,]$month_2_year_0 + ath.aus[ath.aus$name == dir.name,]$month_12_year_0 + ath.aus[ath.aus$name == dir.name,]$month_2_year_2 + ath.aus[ath.aus$name == dir.name,]$month_12_year_2 + ath.aus[ath.aus$name == dir.name,]$full_2years
  # for (file in dir.files[1:2]) {
  #   # read fit file
  #   fitdata <- read.fit(paste0(dir,"/",file))
  #   # check sport type
  #   fitdata.sporttype <- fitdata$session$sport
  #   if (fitdata.sporttype == 2) {
  #     ath.aus[ath.aus$name == dir.name,"n_cycling"] <- ath.aus[ath.aus$name == dir.name,"n_cycling"] + 1
  #   } else if (fitdata.sporttype == 1) {
  #     ath.aus[ath.aus$name == dir.name,"n_running"] <- ath.aus[ath.aus$name == dir.name,"n_running"] + 1
  #   } else {
  #     ath.aus[ath.aus$name == dir.name,"n_other"] <- ath.aus[ath.aus$name == dir.name,"n_other"] + 1
  #   }
  #   # check heart rate if present
  #   if (!is.null(fitdata$record$heart_rate)) {
  #     ath.aus[ath.aus$name == dir.name,"n_heart"] <- ath.aus[ath.aus$name == dir.name,"n_heart"] + 1
  #     ath.aus[ath.aus$name == dir.name,"perc_heart"] <- round(ath.aus[ath.aus$name == dir.name,"n_heart"] / ath.aus[ath.aus$name == dir.name,"total_activities"] * 100,1)
  #   }
  # }
}
```

```{r}
ath.aus
```


*** 

## PRO@HEART in LEUVEN

### Athlete input files


```{r}
# dirs <- list.dirs("~/sgcluster/WORK/BIOINFO/TRAINING_PEAKS/SELECTED_FILES",recursive=F)
dirs_raw.leuven <- list.dirs("PRO_HEART/LEUVEN_PH")
dirs.leuven <- dirs_raw.leuven[str_detect(dirs_raw.leuven,"YEAR")]

# for (dir in dirs){
#   print(dir)
#   print(head(list.files(dir),1))
# }
```

Initialize athlete list

```{r}
ath.leuven<- as.data.frame(unique(do.call(rbind, str_split(dirs.leuven,"/"))[,3]),stringsAsFactors =FALSE)
colnames(ath.leuven) <- "name"
ath.leuven$total_activities <- 0
ath.leuven$month_2_year_0 <- 0
ath.leuven$month_12_year_0 <- 0
ath.leuven$month_2_year_2 <- 0
ath.leuven$month_12_year_2 <- 0
ath.leuven$full_2years <- 0
# ath.leuven$n_cycling <- 0
# ath.leuven$n_running <- 0
# ath.leuven$n_other <- 0
# ath.leuven$n_heart <- 0
# ath.leuven$perc_heart <- 0
```

### Fill list

```{r}
for (dir in dirs.leuven){
  dir.name <- unlist(str_split(dir,"/"))[3]
  dir.month <- tolower(unlist(str_split(dir,"/"))[4])
  dir.files <- list.files(dir,pattern="fit")
  ath.leuven[ath.leuven$name == dir.name,dir.month] <-  length(dir.files)
  ath.leuven[ath.leuven$name == dir.name,"total_activities"] <- ath.leuven[ath.leuven$name == dir.name,]$month_2_year_0 + ath.leuven[ath.leuven$name == dir.name,]$month_12_year_0 + ath.leuven[ath.leuven$name == dir.name,]$month_2_year_2 + ath.leuven[ath.leuven$name == dir.name,]$month_12_year_2 + ath.leuven[ath.leuven$name == dir.name,]$full_2years
  # for (file in dir.files[1:2]) {
  #   # read fit file
  #   fitdata <- read.fit(paste0(dir,"/",file))
  #   # check sport type
  #   fitdata.sporttype <- fitdata$session$sport
  #   if (fitdata.sporttype == 2) {
  #     ath.leuven[ath.leuven$name == dir.name,"n_cycling"] <- ath.leuven[ath.leuven$name == dir.name,"n_cycling"] + 1
  #   } else if (fitdata.sporttype == 1) {
  #     ath.leuven[ath.leuven$name == dir.name,"n_running"] <- ath.leuven[ath.leuven$name == dir.name,"n_running"] + 1
  #   } else {
  #     ath.leuven[ath.leuven$name == dir.name,"n_other"] <- ath.leuven[ath.leuven$name == dir.name,"n_other"] + 1
  #   }
  #   # check heart rate if present
  #   if (!is.null(fitdata$record$heart_rate)) {
  #     ath.leuven[ath.leuven$name == dir.name,"n_heart"] <- ath.leuven[ath.leuven$name == dir.name,"n_heart"] + 1
  #     ath.leuven[ath.leuven$name == dir.name,"perc_heart"] <- round(ath.leuven[ath.leuven$name == dir.name,"n_heart"] / ath.leuven[ath.leuven$name == dir.name,"total_activities"] * 100,1)
  #   }
  # }
}
```

```{r}
ath.leuven
```


*** 
# MASTER @ HEART

## MASTER@HEART in ANTWERPEN

### Athlete input files


```{r}
# dirs <- list.dirs("~/sgcluster/WORK/BIOINFO/TRAINING_PEAKS/SELECTED_FILES",recursive=F)
dirs_raw.ant.mh <- list.dirs("MASTER_HEART/ANTWERPEN_MH")
dirs.ant.mh <- dirs_raw.ant.mh[str_detect(dirs_raw.ant.mh,"MONTH")]
dirs.ant.mh <- dirs.ant.mh[str_detect(dirs.ant.mh,".*/.*/.*/")]

# for (dir in dirs){
#   print(dir)
#   print(head(list.files(dir),1))
# }
```

Initialize athlete list

```{r}
ath.ant.mh<- as.data.frame(unique(do.call(rbind, str_split(dirs.ant.mh,"/"))[,4]),stringsAsFactors =FALSE)
colnames(ath.ant.mh) <- "name"
ath.ant.mh$total_activities <- 0
ath.ant.mh$month_2 <- 0
ath.ant.mh$month_12 <- 0
```

### Fill list

```{r}
for (dir in dirs.ant.mh){
  dir.name <- unlist(str_split(dir,"/"))[4]
  dir.month <- tolower(unlist(str_split(dir,"/"))[3])
  dir.files <- list.files(dir,pattern="fit")
  ath.ant.mh[ath.ant.mh$name == dir.name,dir.month] <-  length(dir.files)
  ath.ant.mh[ath.ant.mh$name == dir.name,"total_activities"] <- ath.ant.mh[ath.ant.mh$name == dir.name,]$month_2 + ath.ant.mh[ath.ant.mh$name == dir.name,]$month_12 
  # for (file in dir.files[1:2]) {
  #   # read fit file
  #   fitdata <- read.fit(paste0(dir,"/",file))
  #   # check sport type
  #   fitdata.sporttype <- fitdata$session$sport
  #   if (fitdata.sporttype == 2) {
  #     ath.ant.mh[ath.ant.mh$name == dir.name,"n_cycling"] <- ath.ant.mh[ath.ant.mh$name == dir.name,"n_cycling"] + 1
  #   } else if (fitdata.sporttype == 1) {
  #     ath.ant.mh[ath.ant.mh$name == dir.name,"n_running"] <- ath.ant.mh[ath.ant.mh$name == dir.name,"n_running"] + 1
  #   } else {
  #     ath.ant.mh[ath.ant.mh$name == dir.name,"n_other"] <- ath.ant.mh[ath.ant.mh$name == dir.name,"n_other"] + 1
  #   }
  #   # check heart rate if present
  #   if (!is.null(fitdata$record$heart_rate)) {
  #     ath.ant.mh[ath.ant.mh$name == dir.name,"n_heart"] <- ath.ant.mh[ath.ant.mh$name == dir.name,"n_heart"] + 1
  #     ath.ant.mh[ath.ant.mh$name == dir.name,"perc_heart"] <- round(ath.ant.mh[ath.ant.mh$name == dir.name,"n_heart"] / ath.ant.mh[ath.ant.mh$name == dir.name,"total_activities"] * 100,1)
  #   }
  # }
}
```

```{r}
ath.ant.mh
```

*** 

## MASTER@HEART in HASSELT

### Athlete input files


```{r}
# dirs <- list.dirs("~/sgcluster/WORK/BIOINFO/TRAINING_PEAKS/SELECTED_FILES",recursive=F)
dirs_raw.hasselt.mh <- list.dirs("MASTER_HEART/HASSELT_MH")
dirs.hasselt.mh <- dirs_raw.hasselt.mh[str_detect(dirs_raw.hasselt.mh,"MONTH")]
dirs.hasselt.mh <- dirs.hasselt.mh[str_detect(dirs.hasselt.mh,".*/.*/.*/")]

# for (dir in dirs){
#   print(dir)
#   print(head(list.files(dir),1))
# }
```

Initialize athlete list

```{r}
ath.hasselt.mh<- as.data.frame(unique(do.call(rbind, str_split(dirs.hasselt.mh,"/"))[,4]),stringsAsFactors =FALSE)
colnames(ath.hasselt.mh) <- "name"
ath.hasselt.mh$total_activities <- 0
ath.hasselt.mh$month_2 <- 0
ath.hasselt.mh$month_12 <- 0
```

### Fill list

```{r}
for (dir in dirs.hasselt.mh){
  dir.name <- unlist(str_split(dir,"/"))[4]
  dir.month <- tolower(unlist(str_split(dir,"/"))[3])
  dir.files <- list.files(dir,pattern="fit")
  ath.hasselt.mh[ath.hasselt.mh$name == dir.name,dir.month] <-  length(dir.files)
  ath.hasselt.mh[ath.hasselt.mh$name == dir.name,"total_activities"] <- ath.hasselt.mh[ath.hasselt.mh$name == dir.name,]$month_2 + ath.hasselt.mh[ath.hasselt.mh$name == dir.name,]$month_12 
  # for (file in dir.files[1:2]) {
  #   # read fit file
  #   fitdata <- read.fit(paste0(dir,"/",file))
  #   # check sport type
  #   fitdata.sporttype <- fitdata$session$sport
  #   if (fitdata.sporttype == 2) {
  #     ath.hasselt.mh[ath.hasselt.mh$name == dir.name,"n_cycling"] <- ath.hasselt.mh[ath.hasselt.mh$name == dir.name,"n_cycling"] + 1
  #   } else if (fitdata.sporttype == 1) {
  #     ath.hasselt.mh[ath.hasselt.mh$name == dir.name,"n_running"] <- ath.hasselt.mh[ath.hasselt.mh$name == dir.name,"n_running"] + 1
  #   } else {
  #     ath.hasselt.mh[ath.hasselt.mh$name == dir.name,"n_other"] <- ath.hasselt.mh[ath.hasselt.mh$name == dir.name,"n_other"] + 1
  #   }
  #   # check heart rate if present
  #   if (!is.null(fitdata$record$heart_rate)) {
  #     ath.hasselt.mh[ath.hasselt.mh$name == dir.name,"n_heart"] <- ath.hasselt.mh[ath.hasselt.mh$name == dir.name,"n_heart"] + 1
  #     ath.hasselt.mh[ath.hasselt.mh$name == dir.name,"perc_heart"] <- round(ath.hasselt.mh[ath.hasselt.mh$name == dir.name,"n_heart"] / ath.hasselt.mh[ath.hasselt.mh$name == dir.name,"total_activities"] * 100,1)
  #   }
  # }
}
```

```{r}
ath.hasselt.mh
```

*** 
## MASTER@HEART in LEUVEN

### Athlete input files


```{r}
# dirs <- list.dirs("~/sgcluster/WORK/BIOINFO/TRAINING_PEAKS/SELECTED_FILES",recursive=F)
dirs_raw.leuven.mh <- list.dirs("MASTER_HEART/LEUVEN_MH")
dirs.leuven.mh <- dirs_raw.leuven.mh[str_detect(dirs_raw.leuven.mh,"MONTH")]
dirs.leuven.mh <- dirs.leuven.mh[str_detect(dirs.leuven.mh,".*/.*/.*/")]

# for (dir in dirs){
#   print(dir)
#   print(head(list.files(dir),1))
# }
```

Initialize athlete list

```{r}
ath.leuven.mh<- as.data.frame(unique(do.call(rbind, str_split(dirs.leuven.mh,"/"))[,4]),stringsAsFactors =FALSE)
colnames(ath.leuven.mh) <- "name"
ath.leuven.mh$total_activities <- 0
ath.leuven.mh$month_2 <- 0
ath.leuven.mh$month_12 <- 0
```

### Fill list

```{r}
for (dir in dirs.leuven.mh){
  dir.name <- unlist(str_split(dir,"/"))[4]
  dir.month <- tolower(unlist(str_split(dir,"/"))[3])
  dir.files <- list.files(dir,pattern="fit")
  ath.leuven.mh[ath.leuven.mh$name == dir.name,dir.month] <-  length(dir.files)
  ath.leuven.mh[ath.leuven.mh$name == dir.name,"total_activities"] <- ath.leuven.mh[ath.leuven.mh$name == dir.name,]$month_2 + ath.leuven.mh[ath.leuven.mh$name == dir.name,]$month_12 
  # for (file in dir.files[1:2]) {
  #   # read fit file
  #   fitdata <- read.fit(paste0(dir,"/",file))
  #   # check sport type
  #   fitdata.sporttype <- fitdata$session$sport
  #   if (fitdata.sporttype == 2) {
  #     ath.leuven.mh[ath.leuven.mh$name == dir.name,"n_cycling"] <- ath.leuven.mh[ath.leuven.mh$name == dir.name,"n_cycling"] + 1
  #   } else if (fitdata.sporttype == 1) {
  #     ath.leuven.mh[ath.leuven.mh$name == dir.name,"n_running"] <- ath.leuven.mh[ath.leuven.mh$name == dir.name,"n_running"] + 1
  #   } else {
  #     ath.leuven.mh[ath.leuven.mh$name == dir.name,"n_other"] <- ath.leuven.mh[ath.leuven.mh$name == dir.name,"n_other"] + 1
  #   }
  #   # check heart rate if present
  #   if (!is.null(fitdata$record$heart_rate)) {
  #     ath.leuven.mh[ath.leuven.mh$name == dir.name,"n_heart"] <- ath.leuven.mh[ath.leuven.mh$name == dir.name,"n_heart"] + 1
  #     ath.leuven.mh[ath.leuven.mh$name == dir.name,"perc_heart"] <- round(ath.leuven.mh[ath.leuven.mh$name == dir.name,"n_heart"] / ath.leuven.mh[ath.leuven.mh$name == dir.name,"total_activities"] * 100,1)
  #   }
  # }
}
```

```{r}
ath.leuven.mh
```

*** 

# TOTAL GROUP

```{r}
total <- rbind(ath.hasselt,ath.leuven,ath.aus)
total <- total[apply(total[,-1], 1, function(row) all(row[-1] > 1 )),]

```


