---
title: "Leah Wright - Lipidomics"
output:
  html_notebook:
    # css:
    # - !expr system.file("rmarkdown/templates/html_vignette/resources/vignette.css",
    #                     package = "rmarkdown")
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    theme: cosmo
    df_print: paged
    highlight: tango
    code_folding: hide
    # fig_width: 12
    # fig_height: 12
# output:
#   epuRate::BAKER:
#     toc: TRUE
#     number_sections: FALSE
#     code_folding: "show"
---

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

***

# Initial setup
## Load libraries

```{r warning=FALSE, results='hide'}
# library(fit)
library(ggplot2)
# library(leaflet)
library(dplyr)
# library(cetcolor)
library(stringr)
library(readxl)
# library(summarytools)
```

## Configure output

```{r setup}
library(knitr)
opts_chunk$set(results = 'asis',      # Can also be set at the chunk-level
               comment = NA,
               prompt  = FALSE,
               cache   = FALSE)
library(summarytools)
st_options(plain.ascii = FALSE,        # Always use this option in Rmd documents
           style        = "rmarkdown", # Always use this option in Rmd documents
           footnote     = NA,          # Makes html-rendered results more concise
           subtitle.emphasis = FALSE)  # Improves layout with some rmardown themes
```

```{r, echo=FALSE}
st_css()
```

***
## Read input

Read file with 2016 and 2017 patients data together:

* Already QCd and mediancentered by lipidomics people
* 2017 individuals lack personal info and CVD measurements


```{r}
file <- "Echo_Lipids no CVD.xlsx"
mydata <- read_xlsx(file)
```

Get only individuals from 2016, as 2017 only have lipidomics data

```{r}
# get only individuals from 2016
data16 <- mydata[1:380,]
```

***

# Data Cleaning
## Preprocess and summarize

Here, I will select and group files in order to get a basic summary of all the dataset.

* Personal info
* Health and disease info
* Cardiac info
* Lipid info (to do separately later)

More details here:
https://cran.r-project.org/web/packages/summarytools/vignettes/Introduction.html



```{r}
fields_to_summarize <- paste(colnames(data16)[1:128],collapse="', '")
fields_nolipids <- c(1:128)
fields_personal <- c(1:24)
fields_health <- c(25:56)
fields_cardiac <- c(57:128)
fields_excluded <- c(1,3,4,5,6,15,32,57:59,64:69,85)
# descr(data16[,fields_n],plain.ascii = F,style="rmarkdown")
```

### Personal info

```{r, warning=F}
fields_personal_OK <- setdiff(fields_personal,fields_excluded)
dfSummary(data16[,fields_personal_OK],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp")
```

### Health and disease info

```{r, warning=F}
fields_health_OK <- setdiff(fields_health,fields_excluded)
dfSummary(data16[,fields_health_OK],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp")
```

### Cardiac info

```{r, warning=F}
fields_cardiac_OK <- setdiff(fields_cardiac,fields_excluded)
dfSummary(data16[,fields_cardiac_OK],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp")
```

```{r}
# descr(data16[,fields_personal_OK],plain.ascii = F,style="rmarkdown")
```

***

## Data by patient

### Missing fields

In this section, I will check missing data for each of the patients. 
We will decide whether we discard some of the patients based on a threshold % of missing values.
For the moment, not checking on lipid values yet.

```{r}
patient.missing.data <- rowSums(is.na(data16[fields_nolipids]))/length(fields_nolipids)*100
# plot(density(patient_missing_data))
# sum(patient_missing_data < 15)
data.breaks <- c(0,5,10,15,20,40,60,100)
data.breaks.chr <- c('(0,5%]','(5,10%]','(10,15%]','(15,20%]','(20,40%]','(40,60%]','(60,100%]')
# cut(patient_missing_data,data.breaks)
# names(split(patient_missing_data,cut(patient_missing_data,data.breaks)))
hist.data <- hist(patient.missing.data,plot=F,breaks=data.breaks)
df.missing.data <- as.data.frame(hist.data$counts)
colnames(df.missing.data) <- "patients"
df.missing.data <- df.missing.data %>%
  mutate(
    missing = data.breaks.chr,
    perc.patients = round(patients/length(patient.missing.data)*100,2),
    cum.patients = cumsum(patients),
    cum.perc = cumsum(perc.patients)
  ) %>%
  select(
    missing,patients,cum.patients,perc.patients,cum.perc,
  )
kable(df.missing.data)
```

So, if we discarted all patients with **more than 15%** missing data, we would only lose 4% of them.

An example of >40% missing data (all data and missing fields), looks like patient missed the ultrasound test:

```{r}
data16[1,fields_nolipids]
data16[1,fields_nolipids][as.logical(is.na(data16[1,fields_nolipids]))]
```
And ~10%, missing some values (common with most patients):

```{r}
data16[2,fields_nolipids]
data16[2,fields_nolipids][as.logical(is.na(data16[2,fields_nolipids]))]
```

# Lipidomics analysis

## Preprocess and cleaning

For a first data check, I will keep using 2016 individuals, without removing any of the high % missing data as shown above.

```{r}
fields_lipids <- c(1,129:743)
lipids <- data16[fields_lipids]
dfSummary(lipids[,1:4],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp")
```

11 cases don't have lipidomics data (doublechecked in excel), will be discarded.

```{r}
lipid_mask <- !is.na(lipids$Sample)
lipids <- lipids[lipid_mask,]
lipids.metadata <- data16[lipid_mask, fields_nolipids]
dfSummary(lipids[,c(1:4,616)],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp")
```

Lipid PC(O-38:5) is also NULL, will be removed.

```{r}
colnames(lipids)[ apply(lipids, 2, anyNA) ]
which( colnames (lipids) == colnames(lipids)[ apply(lipids, 2, anyNA) ])
lipids <- lipids [,-354]
```

Get lipids table as in sample ID vs all lipid values

```{r}
lipids <- lipids [,c(-1,-2,-3)]
dim(lipids)
```

**Final set of 369 patients with 612 lipids.**

## PCA

Check with PCA

Log2 transform first

```{r}
lipids <- log2(lipids)
```


```{r}
lipids.pca <- prcomp(lipids,center=TRUE,scale. = T)
pc1var <- round(summary(lipids.pca)$importance[2,1]*100,1)
pc2var <- round(summary(lipids.pca)$importance[2,2]*100,1)
# lipids.pca <- prcomp(lipids[,2:10],center=TRUE, scale. = TRUE)
# head(summary(lipids.pca))
```

```{r, echo = F, fig.width=10}
par(mfrow=c(2,3))
par(oma = c(4, 4, 0.3, 0)) # make room (i.e. the 4's) for the overall x and y axis titles
par(mar = c(2, 2, 2, 1)) # make the plots be closer together
mycex.text=0.7
mycex.legend=0.8
mycex.plot=1.2
# plot 1
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=lipids.metadata$site_no,
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="site_name",cex=mycex.plot)
legend("topright",fill=unique(lipids.metadata$site_no),legend=c("Alice Springs", "Palm Island"),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# plot 2
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=lipids.metadata$sex+1,
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="sex",cex=mycex.plot)
legend("topright",fill=unique(lipids.metadata$sex+1),legend=c("Male", "Female"),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# plot 3
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=lipids.metadata$agecateg,
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="agecateg",cex=mycex.plot)
legend("topright",fill=sort(unique(lipids.metadata$agecateg)),
       legend=c("< 35", "35 < 44","45 < 54","> 55"),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# plot 4
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=as.factor(lipids.metadata$BMIClassif),
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="BMI",cex=mycex.plot)
legend("topright",fill=unique(as.factor(lipids.metadata$BMIClassif)),
       legend=paste(unique(as.factor(lipids.metadata$BMIClassif))),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# plot 5
sel_var <- lipids.metadata$SBPClassif
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=as.factor(sel_var),
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="BPClassif",cex=mycex.plot)
legend("topright",fill=unique(as.factor(sel_var)),
       legend=paste(unique(as.factor(sel_var))),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# plot 6
sel_var <- lipids.metadata$diabetes
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=as.factor(sel_var),
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="diabetes",cex=mycex.plot)
legend("topright",fill=unique(as.factor(sel_var)),
       legend=paste(unique(as.factor(sel_var))),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# axis text
mtext(paste0("PC1 (",pc1var,"% explained var.)"), side = 1, outer = TRUE, line = 2)
mtext(paste0("PC2 (",pc2var,"% explained var.)"), side = 2, outer = TRUE, line = 2)
```



```{r}
# library(ggbiplot)
# ggbiplot(lipids.pca,labels=rownames(lipids))
# ggbiplot(lipids.pca)
# biplot(lipids.pca)
```

## Heatmap

```{r}
library(gplots)
mydata <- cor(lipids, method = "pearson", use = "complete.obs")
breaks = seq(-1, 1, length.out=50)
gradient1 = colorpanel( sum( breaks[-1]<=-0.5 ), "yellow", "green" )
gradient2 = colorpanel( sum( breaks[-1]>-0.5 & breaks[-1]<=0 ), "green", "blue" )
gradient3 = colorpanel( sum( breaks[-1]>0 & breaks[-1]<=0.5 ), "blue", "darkblue" )
gradient4 = colorpanel( sum( breaks[-1]>0.5 ), "darkblue", "black" )
hm.colors = c(gradient1, gradient2, gradient3, gradient4)
heatmap.2(mydata,dendrogram='row',Rowv=T,Colv=T,trace='none',col=hm.colors,breaks=breaks,
          labRow = F, labCol = F,margins = c(2, 2),
          xlab="Lipid Correlation")

#FULL matrix
# mydataFull <- t(mydata)
# mydataFull[lower.tri(mydataFull)]<-mydata[lower.tri(mydata)]
# heatmap.2(mat,trace='none',col=col)

```


