---
title: "Leah Wright - Lipidomics"
author: "Baker Bioinformatics"
date: "`r format(Sys.time(), '%d %B %Y')`"
# output:
  # html_notebook:
  #   # css:
  #   # - !expr system.file("rmarkdown/templates/html_vignette/resources/vignette.css",
  #   #                     package = "rmarkdown")
  #   toc: true
  #   toc_depth: 3
  #   toc_float:
  #     collapsed: false
  #     smooth_scroll: true
  #   theme: cosmo
  #   df_print: paged
  #   highlight: tango
  #   code_folding: show
  #   # fig_width: 12
  #   # fig_height: 12
output:
  epuRate::BAKER:
    toc: TRUE
    number_sections: FALSE
    code_folding: "show"
---

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

***

# Initial setup
## Load libraries

```{r warning=FALSE, results='hide'}
# library(fit)
library(ggplot2)
# library(leaflet)
library(dplyr)
# library(cetcolor)
library(stringr)
library(readxl)
# library(summarytools)
```

## Configure output

```{r setup}
library(knitr)
opts_chunk$set(results = 'asis',      # Can also be set at the chunk-level
               comment = NA,
               prompt  = FALSE,
               cache   = FALSE)
library(summarytools)
st_options(plain.ascii = FALSE,        # Always use this option in Rmd documents
           style        = "rmarkdown", # Always use this option in Rmd documents
           footnote     = NA,          # Makes html-rendered results more concise
           subtitle.emphasis = FALSE,  # Improves layout with some rmardown themes
           # bootstrap.css = FALSE,
)
```

```{r, echo=TRUE}
st_css()
```

***
# Read input

Read file with 2016 and 2017 patients data together:

* Already QCd and mediancentered by lipidomics people
* 2017 individuals lack personal info and CVD measurements
  * __Corrected in the last update sent by Leah__
  
```{r}
# file <- "Echo_Lipids no CVD.xlsx"
file <- "final_dataset.xlsx"
mydata <- read_xlsx(file)
```

***

# Data Cleaning
## Preprocess and summarize

Here, we will select and group files in order to get a basic summary of all the dataset.

* Personal info
* Health and disease info
* Cardiac info
* Lipid info (to do separately later)

We will use these results to sort and filter out the patients without enough data in each of the groups (e.g. _not enough cardiac info or lack of health measurements_)

More details about the summarytools library here:  
https://cran.r-project.org/web/packages/summarytools/vignettes/Introduction.html

Divide all fields in the different groups:

```{r}
# fields_to_summarize <- paste(colnames(mydata)[1:128],collapse="', '")
fields_nolipids <- c(1:101)
fields_lipids <- c(1,102:716)
fields_personal <- c(1:23)
fields_health <- c(24:53)
fields_cardiac <- c(54:101)
# # exclude uninformative fields for the report, such as sno 1, site_name 2, date_enrol 4, pt_initials 5, dob 6, repeated ID 54
# fields_excluded <- c(1,2,4,5,6,54)
fields_excluded <- NULL
```

### Personal info

Corrections:

* Change pregnant == 2 (not-pregnant), to 0, as it makes more sense.
  * Men will still have NA and women will have 0 or 1.
  * Add waist to hip ratio

```{r}
mydata$pregnant[mydata$pregnant == 2] <- 0
mydata <- mydata %>% mutate(waist_hip_ratio = waist/hip)
```

Summary:

```{r message=FALSE}
fields_personal_OK <- setdiff(fields_personal,fields_excluded)
dfSummary(mydata$waist_hip_ratio,style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp")
print(dfSummary(mydata$waist_hip_ratio,style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp"),method='render')
```
