---
title: "Leah Wright - Lipidomics"
author: "[Baker Bioinformatics](https://intranet.baker.edu.au/research-resources/platform-technologies/bioinformatics-program/)"
date: "`r format(Sys.time(), '%d %B %Y')`"
mail: "bioinformatics@baker.edu.au"
# output:
#   html_notebook:
#     # css:
#     # - !expr system.file("rmarkdown/templates/html_vignette/resources/vignette.css",
#     #                     package = "rmarkdown")
#     toc: true
#     toc_depth: 3
#     toc_float:
#       collapsed: false
#       smooth_scroll: true
#     theme: cosmo
#     df_print: paged
#     highlight: tango
#     code_folding: show
#     # fig_width: 12
#     # fig_height: 12
output:
  epuRate::BAKER:
    toc: TRUE
    number_sections: FALSE
    code_folding: "show"
---

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

***

# Initial setup
## Load libraries

```{r warning=FALSE, results='hide'}
# library(fit)
library(ggplot2)
# library(leaflet)
library(dplyr)
# library(cetcolor)
library(stringr)
library(readxl)
# library(summarytools)
```

## Configure output

```{r setup}
library(knitr)
opts_chunk$set(results = 'asis',      # Can also be set at the chunk-level
               comment = NA,
               prompt  = FALSE,
               cache   = FALSE)
library(summarytools)
st_options(plain.ascii = FALSE,        # Always use this option in Rmd documents
           style        = "rmarkdown", # Always use this option in Rmd documents
           footnote     = NA,          # Makes html-rendered results more concise
           subtitle.emphasis = FALSE)  # Improves layout with some rmardown themes
```

```{r, echo=FALSE}
st_css()
```

***
# Read input

Read file with 2016 and 2017 patients data together:

* Already QCd and mediancentered by lipidomics people
* 2017 individuals lack personal info and CVD measurements
  * __Corrected in the last update sent by Leah__
  
```{r}
# file <- "Echo_Lipids no CVD.xlsx"
file <- "final_dataset.xlsx"
mydata <- read_xlsx(file)
```

***

# Data Cleaning
## Preprocess and summarize

Here, we will select and group files in order to get a basic summary of all the dataset.

* Personal info
* Health and disease info
* Cardiac info
* Lipid info (to do separately later)

We will use these results to sort and filter out the patients without enough data in each of the groups (e.g. _not enough cardiac info or lack of health measurements_)

More details about the summarytools library here:  
https://cran.r-project.org/web/packages/summarytools/vignettes/Introduction.html

Divide all fields in the different groups:

```{r}
# fields_to_summarize <- paste(colnames(mydata)[1:128],collapse="', '")
fields_nolipids <- c(1:101)
fields_lipids <- c(1,102:716)
fields_personal <- c(1:23)
fields_health <- c(24:53)
fields_cardiac <- c(54:101)
# # exclude uninformative fields for the report, such as sno 1, site_name 2, date_enrol 4, pt_initials 5, dob 6, repeated ID 54
# fields_excluded <- c(1,2,4,5,6,54)
fields_excluded <- NULL
```

### Personal info

Corrections:

* Change pregnant == 2 (not-pregnant), to 0, as it makes more sense.
  * Men will still have NA and women will have 0 or 1.
  * Add waist to hip ratio

```{r}
mydata$pregnant[mydata$pregnant == 2] <- 0
mydata <- mydata %>% mutate(waist_hip_ratio = waist/hip)
```

Summary:

```{r, warning=F}
fields_personal_OK <- setdiff(fields_personal,fields_excluded)
print(dfSummary(mydata[,fields_personal_OK],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp"),method='render')
print(dfSummary(mydata$waist_hip_ratio,style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp"),method='render')
```

### Health and disease info

Some columns have missing data but does not appear critical:

* Some blood pressure values and category classifications (1-3%).
  * We will use average SBP (mainly) and DBP which are complete.
* Alc_quantity (25% missing), but will keep as Leah wants to include it when we have the data.
  * Refers to quantity alcohol column in 1:day, 2:week and 3:month.
  * Would it make sense to standardize all to month? Considering the missing values here as if they were month?
* ldlmmolL has 28 missing values (5.73%), we will keep it and check when analysing later on.

Removing patients with >% of missing data might decrease the amount of column missingness.

```{r, warning=F}
fields_health_OK <- setdiff(fields_health,fields_excluded)
print(dfSummary(mydata[,fields_health_OK],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp"),method='render')
```

### Cardiac info

Some columns have missing data but we will keep them as Leah wants it to be included:

* Remove column with ID (54), as it's a repetition of column 5.
* ValveReplacement and the following columns have high missing data, but will keep them.
* Special focus on columns GLS_Peak, Esepta, Elat, eavg, Eemed, EELat, EEavg, LA_Index and LVMass_Index, as they are the most interesting ones for Leah's analysis.
  * All of them have low missingness.

As section before, removing patients with >% of missing data might decrease the amount of column missingness.


```{r, warning=F}
fields_cardiac_OK <- setdiff(fields_cardiac,fields_excluded)
print(dfSummary(mydata[,fields_cardiac_OK],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp"),method='render')
```

***

## Data by patient

### Missing fields

In this section, we will check missing data for each of the patients. 

* We will decide whether we discard some of the patients based on a threshold % of missing values.
* For the moment, not checking on lipid values yet.

```{r}
patient.missing.data <- rowSums(is.na(mydata[fields_nolipids]))/length(fields_nolipids)*100
data.breaks <- c(0,5,10,15,20,40,60,100)
data.breaks.chr <- c('(0,5%]','(5,10%]','(10,15%]','(15,20%]','(20,40%]','(40,60%]','(60,100%]')
hist.data <- hist(patient.missing.data,plot=F,breaks=data.breaks)
df.missing.data <- as.data.frame(hist.data$counts)
colnames(df.missing.data) <- "patients"
df.missing.data <- df.missing.data %>%
  mutate(
    missing = data.breaks.chr,
    perc.patients = round(patients/length(patient.missing.data)*100,2),
    cum.patients = cumsum(patients),
    cum.perc = cumsum(perc.patients)
  ) %>%
  select(
    missing,patients,cum.patients,perc.patients,cum.perc,
  )
df.missing.data
```

So, if we discarted all patients with **more than 15%** missing data, we would only lose 2% of them.

An example of >40% missing data (all data and missing fields), looks like patient missed the ultrasound test:

```{r}
mydata[1,fields_nolipids]
mydata[1,fields_nolipids][as.logical(is.na(mydata[1,fields_nolipids]))]
```

And an example of ~10%, missing some values (common with most patients):

```{r}
mydata[11,fields_nolipids]
mydata[11,fields_nolipids][as.logical(is.na(mydata[11,fields_nolipids]))]
```

## Final clean set

To summarize:

* Remove individuals with >15% missing variables
* Get clean data with only no-lipid variables (including W2H ratio and removing repeated ID)

```{r}
dim(mydata)
mydata.clean <- mydata[patient.missing.data < 15,]
dim(mydata.clean)
# get nolipids and W2H Ratio
mydata.clean.metadata <- mydata.clean[,c(fields_nolipids,717)]
dim(mydata.clean.metadata)
# remove repeated ID
mydata.clean.metadata$ID <- NULL
dim(mydata.clean.metadata)
```

* Summarize again all the variables and see missingness:
  * No big changes, but a bit of cleanup on some variables

```{r message=FALSE, warning=FALSE}
# colSums(is.na(mydata.clean.metadata))
# descr(mydata.clean.metadata, stats = c("mean", "min", "max","n.valid"),headings=F,plain.ascii = F,style="rmarkdown")
missingness.summary <- round(t(as.data.frame(colSums(is.na(mydata.clean.metadata))))/dim(mydata.clean.metadata)[1]*100,2)
original.data <- mydata[,c(fields_nolipids,717)]
original.data$ID <- NULL
missingness.summary <- rbind(missingness.summary,round(t(as.data.frame(colSums(is.na(original.data))))/dim(original.data)[1]*100,2))
rownames(missingness.summary) <- c("Miss. clean data (%)","Miss. original data (%)")
as.data.frame(missingness.summary)
```

Clean data with __480 individuals and 101 variables__, containing personal, health and disease info and cardiac measurements.

***

# Lipidomics analysis

## Preprocess and cleaning

For a first data check, we will keep using the 480 individuals from the clean dataset above.

```{r}
fields_lipids <- c(1,102:716)
lipids <- mydata.clean[fields_lipids]
print(dfSummary(lipids[,1:4],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp"),method='render')
```

19 of the remaining patients don't have lipidomics data (doublechecked in excel), will be discarded.

```{r}
lipid_mask <- !is.na(lipids$Sample)
lipids <- lipids[lipid_mask,]
print(dfSummary(lipids[,c(1:4,616)],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp"),method='render')
```

Remove same patients from clean dataset (metadata)

```{r}
mydata.clean.metadata <- mydata.clean.metadata[lipid_mask,]
dim(mydata.clean.metadata)
dim(lipids)
```


Lipid PC(O-38:5) is also NULL, will be removed.

```{r}
colnames(lipids)[ apply(lipids, 2, anyNA) ]
which( colnames (lipids) == colnames(lipids)[ apply(lipids, 2, anyNA) ])
lipids <- lipids [,-which( colnames (lipids) == colnames(lipids)[ apply(lipids, 2, anyNA) ])]
dim(lipids)
```

Get lipids table as in sample ID (sno) in rows vs all lipid values in columns

```{r}
library(tibble)
# remove 'Sample' and 'sample details' columns convert to dataframe with sno as rownames
lipids <- lipids [,c(-2,-3)]
lipids <- column_to_rownames(lipids,var="sno")
dim(lipids)
```

Save clean dataset in excel format

```{r}
# library(xlsx)
# clean.dataset <- cbind(mydata.clean.metadata,lipids)
# write.xlsx(clean.dataset,file="clean_dataset.xlsx")
# write.csv(clean.dataset,file="clean_dataset.csv")
```


**Final set of 461 patients with 612 lipids.**


## PCA

Check with PCA

Log2 transform first

```{r}
lipids <- log2(lipids)
lipids.metadata <- mydata.clean.metadata
```


```{r}
lipids.pca <- prcomp(lipids,center=TRUE,scale. = T)
pc1var <- round(summary(lipids.pca)$importance[2,1]*100,1)
pc2var <- round(summary(lipids.pca)$importance[2,2]*100,1)
# lipids.pca <- prcomp(lipids[,2:10],center=TRUE, scale. = TRUE)
# head(summary(lipids.pca))
```

```{r, echo = F, fig.width=10}
par(mfrow=c(2,3))
par(oma = c(4, 4, 0.3, 0)) # make room (i.e. the 4's) for the overall x and y axis titles
par(mar = c(2, 2, 2, 1)) # make the plots be closer together
mycex.text=0.7
mycex.legend=0.8
mycex.plot=1.2
# plot 1
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=lipids.metadata$site_no,
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="site_name",cex=mycex.plot)
legend("topright",fill=unique(lipids.metadata$site_no),legend=c("Alice Springs", "Palm Island"),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# plot 2
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=lipids.metadata$sex+1,
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="sex",cex=mycex.plot)
legend("topright",fill=unique(lipids.metadata$sex+1),legend=c("Male", "Female"),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# plot 3
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=lipids.metadata$agecateg,
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="agecateg",cex=mycex.plot)
legend("topright",fill=sort(unique(lipids.metadata$agecateg)),
       legend=c("< 35", "35 < 44","45 < 54","> 55"),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# plot 4
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=as.factor(lipids.metadata$BMIClassif),
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="BMI",cex=mycex.plot)
legend("topright",fill=unique(as.factor(lipids.metadata$BMIClassif)),
       legend=paste(unique(as.factor(lipids.metadata$BMIClassif))),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# plot 5
sel_var <- lipids.metadata$SBPClassif
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=as.factor(sel_var),
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="BPClassif",cex=mycex.plot)
legend("topright",fill=unique(as.factor(sel_var)),
       legend=paste(unique(as.factor(sel_var))),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# plot 6
sel_var <- lipids.metadata$diabetes
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=as.factor(sel_var),
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="diabetes",cex=mycex.plot)
legend("topright",fill=unique(as.factor(sel_var)),
       legend=paste(unique(as.factor(sel_var))),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# axis text
mtext(paste0("PC1 (",pc1var,"% explained var.)"), side = 1, outer = TRUE, line = 2)
mtext(paste0("PC2 (",pc2var,"% explained var.)"), side = 2, outer = TRUE, line = 2)
```



```{r}
# library(ggbiplot)
# ggbiplot(lipids.pca,labels=rownames(lipids))
# ggbiplot(lipids.pca)
# biplot(lipids.pca)
```

## Heatmap

```{r}
library(gplots)
m <- cor(lipids, method = "pearson", use = "complete.obs")
breaks = seq(-1, 1, length.out=50)
gradient1 = colorpanel( sum( breaks[-1]<=-0.5 ), "yellow", "green" )
gradient2 = colorpanel( sum( breaks[-1]>-0.5 & breaks[-1]<=0 ), "green", "blue" )
gradient3 = colorpanel( sum( breaks[-1]>0 & breaks[-1]<=0.5 ), "blue", "darkblue" )
gradient4 = colorpanel( sum( breaks[-1]>0.5 ), "darkblue", "black" )
hm.colors = c(gradient1, gradient2, gradient3, gradient4)
heatmap.2(m,dendrogram='row',Rowv=T,Colv=T,trace='none',col=hm.colors,breaks=breaks,
          labRow = F, labCol = F,margins = c(2, 2),
          xlab="Lipid Correlation")

#FULL matrix
# mydataFull <- t(mydata)
# mydataFull[lower.tri(mydataFull)]<-mydata[lower.tri(mydata)]
# heatmap.2(mat,trace='none',col=col)

```


