---
title: "Leah Wright - Lipidomics"
author: "[Baker Bioinformatics (Artika and Sergio)](https://intranet.baker.edu.au/research-resources/platform-technologies/bioinformatics-program/)"
date: "`r format(Sys.time(), '%d %B %Y')`"
mail: "bioinformatics@baker.edu.au"
# output:
#   html_notebook:
#     # css:
#     # - !expr system.file("rmarkdown/templates/html_vignette/resources/vignette.css",
#     #                     package = "rmarkdown")
#     toc: true
#     toc_depth: 3
#     toc_float:
#       collapsed: false
#       smooth_scroll: true
#     theme: cosmo
#     df_print: paged
#     highlight: tango
#     code_folding: show
#     # fig_width: 12
#     # fig_height: 12
output:
  epuRate::BAKER:
    toc: TRUE
    number_sections: FALSE
    code_folding: "show"
---

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

***

# Initial setup
## Load libraries

```{r warning=FALSE, results='hide'}
# library(fit)
library(ggplot2)
# library(leaflet)
library(dplyr)
# library(cetcolor)
library(stringr)
library(readxl)
# library(summarytools)
# 2nd part
library(rmarkdown)    # install.packages("rmarkdown")
library(tidyr)
library(epuRate)      # devtools::install_github("holtzy/epuRate", force=TRUE)
```

## Configure output

```{r setup}
library(knitr)
opts_chunk$set(results = 'asis',      # Can also be set at the chunk-level
               comment = NA,
               prompt  = FALSE,
               cache   = FALSE)
library(summarytools)
st_options(plain.ascii = FALSE,        # Always use this option in Rmd documents
           style        = "rmarkdown", # Always use this option in Rmd documents
           footnote     = NA,          # Makes html-rendered results more concise
           subtitle.emphasis = FALSE)  # Improves layout with some rmardown themes
```

```{r, echo=FALSE}
st_css()
```

***
# Read input

Read file with 2016 and 2017 patients data together:

* Already QCd and mediancentered by lipidomics people
* 2017 individuals lack personal info and CVD measurements
  * __Corrected in the last update sent by Leah__
  
```{r}
# file <- "Echo_Lipids no CVD.xlsx"
file <- "final_dataset.xlsx"
mydata <- read_xlsx(file)
```

***

# Data Cleaning
## Preprocess and summarize

Here, we will select and group files in order to get a basic summary of all the dataset.

* Personal info
* Health and disease info
* Cardiac info
* Lipid info (to do separately later)

We will use these results to sort and filter out the patients without enough data in each of the groups (e.g. _not enough cardiac info or lack of health measurements_)

More details about the summarytools library here:  
https://cran.r-project.org/web/packages/summarytools/vignettes/Introduction.html

Divide all fields in the different groups:

```{r}
# fields_to_summarize <- paste(colnames(mydata)[1:128],collapse="', '")
fields_nolipids <- c(1:101)
fields_lipids <- c(1,102:716)
fields_personal <- c(1:23)
fields_health <- c(24:53)
fields_cardiac <- c(54:101)
# # exclude uninformative fields for the report, such as sno 1, site_name 2, date_enrol 4, pt_initials 5, dob 6, repeated ID 54
# fields_excluded <- c(1,2,4,5,6,54)
fields_excluded <- NULL
```

### Personal info

Corrections:

* Change pregnant == 2 (not-pregnant), to 0, as it makes more sense.
  * Men will still have NA and women will have 0 or 1.
  * Add waist to hip ratio

```{r}
mydata$pregnant[mydata$pregnant == 2] <- 0
mydata <- mydata %>% mutate(waist_hip_ratio = waist/hip)
```

Summary:

```{r, warning=F}
fields_personal_OK <- setdiff(fields_personal,fields_excluded)
print(dfSummary(mydata[,fields_personal_OK],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp"),method='render')
print(dfSummary(mydata$waist_hip_ratio,style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp"),method='render')
```

### Health and disease info

Some columns have missing data but does not appear critical:

* Some blood pressure values and category classifications (1-3%).
  * We will use average SBP (mainly) and DBP which are complete.
* Alc_quantity (25% missing), but will keep as Leah wants to include it when we have the data.
  * Refers to quantity alcohol column in 1:day, 2:week and 3:month.
  * Would it make sense to standardize all to month? Considering the missing values here as if they were month?
* ldlmmolL has 28 missing values (5.73%), we will keep it and check when analysing later on.

Removing patients with >% of missing data might decrease the amount of column missingness.

```{r, warning=F}
fields_health_OK <- setdiff(fields_health,fields_excluded)
print(dfSummary(mydata[,fields_health_OK],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp"),method='render')
```

### Cardiac info

Some columns have missing data but we will keep them as Leah wants it to be included:

* Remove column with ID (54), as it's a repetition of column 5.
* ValveReplacement and the following columns have high missing data, but will keep them.
* Special focus on columns GLS_Peak, Esepta, Elat, eavg, Eemed, EELat, EEavg, LA_Index and LVMass_Index, as they are the most interesting ones for Leah's analysis.
  * All of them have low missingness.

As section before, removing patients with >% of missing data might decrease the amount of column missingness.


```{r, warning=F}
fields_cardiac_OK <- setdiff(fields_cardiac,fields_excluded)
print(dfSummary(mydata[,fields_cardiac_OK],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp"),method='render')
```

***

## Data by patient

### Missing fields

In this section, we will check missing data for each of the patients. 

* We will decide whether we discard some of the patients based on a threshold % of missing values.
* For the moment, not checking on lipid values yet.

```{r}
patient.missing.data <- rowSums(is.na(mydata[fields_nolipids]))/length(fields_nolipids)*100
data.breaks <- c(0,5,10,15,20,40,60,100)
data.breaks.chr <- c('(0,5%]','(5,10%]','(10,15%]','(15,20%]','(20,40%]','(40,60%]','(60,100%]')
hist.data <- hist(patient.missing.data,plot=F,breaks=data.breaks)
df.missing.data <- as.data.frame(hist.data$counts)
colnames(df.missing.data) <- "patients"
df.missing.data <- df.missing.data %>%
  mutate(
    missing = data.breaks.chr,
    perc.patients = round(patients/length(patient.missing.data)*100,2),
    cum.patients = cumsum(patients),
    cum.perc = cumsum(perc.patients)
  ) %>%
  select(
    missing,patients,cum.patients,perc.patients,cum.perc,
  )
df.missing.data
```

So, if we discarted all patients with **more than 15%** missing data, we would only lose 2% of them.

An example of >40% missing data (all data and missing fields), looks like patient missed the ultrasound test:

```{r}
mydata[1,fields_nolipids]
mydata[1,fields_nolipids][as.logical(is.na(mydata[1,fields_nolipids]))]
```

And an example of ~10%, missing some values (common with most patients):

```{r}
mydata[11,fields_nolipids]
mydata[11,fields_nolipids][as.logical(is.na(mydata[11,fields_nolipids]))]
```

## Final clean set

To summarize:

* Remove individuals with >15% missing variables
* Get clean data with only no-lipid variables (including W2H ratio and removing repeated ID)

```{r}
dim(mydata)
mydata.clean <- mydata[patient.missing.data < 15,]
dim(mydata.clean)
# get nolipids and W2H Ratio
mydata.clean.metadata <- mydata.clean[,c(fields_nolipids,717)]
dim(mydata.clean.metadata)
# remove repeated ID
mydata.clean.metadata$ID <- NULL
dim(mydata.clean.metadata)
```

* Summarize again all the variables and see missingness:
  * No big changes, but a bit of cleanup on some variables

```{r message=FALSE, warning=FALSE}
# colSums(is.na(mydata.clean.metadata))
# descr(mydata.clean.metadata, stats = c("mean", "min", "max","n.valid"),headings=F,plain.ascii = F,style="rmarkdown")
missingness.summary <- round(t(as.data.frame(colSums(is.na(mydata.clean.metadata))))/dim(mydata.clean.metadata)[1]*100,2)
original.data <- mydata[,c(fields_nolipids,717)]
original.data$ID <- NULL
missingness.summary <- rbind(missingness.summary,round(t(as.data.frame(colSums(is.na(original.data))))/dim(original.data)[1]*100,2))
rownames(missingness.summary) <- c("Miss. clean data (%)","Miss. original data (%)")
as.data.frame(missingness.summary)
```

Clean data with __480 individuals and 101 variables__, containing personal, health and disease info and cardiac measurements.

***

# Lipidomics Preliminar analysis

## Preprocess and cleaning

For a first data check, we will keep using the 480 individuals from the clean dataset above.

```{r}
fields_lipids <- c(1,102:716)
lipids <- mydata.clean[fields_lipids]
print(dfSummary(lipids[,1:4],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp"),method='render')
```

19 of the remaining patients don't have lipidomics data (doublechecked in excel), will be discarded.

```{r}
lipid_mask <- !is.na(lipids$Sample)
lipids <- lipids[lipid_mask,]
print(dfSummary(lipids[,c(1:4,616)],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp"),method='render')
```

Remove same patients from clean dataset (metadata)

```{r}
mydata.clean.metadata <- mydata.clean.metadata[lipid_mask,]
dim(mydata.clean.metadata)
dim(lipids)
```


Lipid PC(O-38:5) is also NULL, will be removed.

```{r}
colnames(lipids)[ apply(lipids, 2, anyNA) ]
which( colnames (lipids) == colnames(lipids)[ apply(lipids, 2, anyNA) ])
lipids <- lipids [,-which( colnames (lipids) == colnames(lipids)[ apply(lipids, 2, anyNA) ])]
dim(lipids)
```

Get lipids table as in sample ID (sno) in rows vs all lipid values in columns

```{r}
library(tibble)
# remove 'Sample' and 'sample details' columns convert to dataframe with sno as rownames
lipids <- lipids [,c(-2,-3)]
lipids <- column_to_rownames(lipids,var="sno")
dim(lipids)
```

Save clean dataset in excel format

```{r}
# library(xlsx)
# clean.dataset <- cbind(mydata.clean.metadata,lipids)
# write.xlsx(clean.dataset,file="clean_dataset.xlsx")
# write.csv(clean.dataset,file="clean_dataset.csv")
```


**Final set of 461 patients with 612 lipids.**


## PCA

Check with PCA

Log2 transform first

```{r}
lipids <- log2(lipids)
lipids.metadata <- mydata.clean.metadata
```


```{r}
lipids.pca <- prcomp(lipids,center=TRUE,scale. = T)
pc1var <- round(summary(lipids.pca)$importance[2,1]*100,1)
pc2var <- round(summary(lipids.pca)$importance[2,2]*100,1)
# lipids.pca <- prcomp(lipids[,2:10],center=TRUE, scale. = TRUE)
# head(summary(lipids.pca))
```

```{r, echo = F, fig.width=10}
par(mfrow=c(2,3))
par(oma = c(4, 4, 0.3, 0)) # make room (i.e. the 4's) for the overall x and y axis titles
par(mar = c(2, 2, 2, 1)) # make the plots be closer together
mycex.text=0.7
mycex.legend=0.8
mycex.plot=1.2
# plot 1
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=lipids.metadata$site_no,
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="site_name",cex=mycex.plot)
legend("topright",fill=unique(lipids.metadata$site_no),legend=c("Alice Springs", "Palm Island"),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# plot 2
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=lipids.metadata$sex+1,
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="sex",cex=mycex.plot)
legend("topright",fill=unique(lipids.metadata$sex+1),legend=c("Male", "Female"),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# plot 3
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=lipids.metadata$agecateg,
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="agecateg",cex=mycex.plot)
legend("topright",fill=sort(unique(lipids.metadata$agecateg)),
       legend=c("< 35", "35 < 44","45 < 54","> 55"),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# plot 4
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=as.factor(lipids.metadata$BMIClassif),
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="BMI",cex=mycex.plot)
legend("topright",fill=unique(as.factor(lipids.metadata$BMIClassif)),
       legend=paste(unique(as.factor(lipids.metadata$BMIClassif))),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# plot 5
sel_var <- lipids.metadata$SBPClassif
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=as.factor(sel_var),
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="BPClassif",cex=mycex.plot)
legend("topright",fill=unique(as.factor(sel_var)),
       legend=paste(unique(as.factor(sel_var))),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# plot 6
sel_var <- lipids.metadata$diabetes
plot(lipids.pca$x[,1],lipids.pca$x[,2],
     pch=16,col=as.factor(sel_var),
     xlab=paste0("PC1 (",pc1var,"% explained var.)"),
     ylab=paste0("PC2 (",pc2var,"% explained var.)"),
     main="diabetes",cex=mycex.plot)
legend("topright",fill=unique(as.factor(sel_var)),
       legend=paste(unique(as.factor(sel_var))),cex=mycex.legend)
text(lipids.pca$x[,1]+1.5,lipids.pca$x[,2],labels = lipids.metadata$sno,cex=mycex.text)
# axis text
mtext(paste0("PC1 (",pc1var,"% explained var.)"), side = 1, outer = TRUE, line = 2)
mtext(paste0("PC2 (",pc2var,"% explained var.)"), side = 2, outer = TRUE, line = 2)
```



```{r}
# library(ggbiplot)
# ggbiplot(lipids.pca,labels=rownames(lipids))
# ggbiplot(lipids.pca)
# biplot(lipids.pca)
```

## Heatmap

```{r}
library(gplots)
m <- cor(lipids, method = "pearson", use = "complete.obs")
breaks = seq(-1, 1, length.out=50)
gradient1 = colorpanel( sum( breaks[-1]<=-0.5 ), "yellow", "green" )
gradient2 = colorpanel( sum( breaks[-1]>-0.5 & breaks[-1]<=0 ), "green", "blue" )
gradient3 = colorpanel( sum( breaks[-1]>0 & breaks[-1]<=0.5 ), "blue", "darkblue" )
gradient4 = colorpanel( sum( breaks[-1]>0.5 ), "darkblue", "black" )
hm.colors = c(gradient1, gradient2, gradient3, gradient4)
heatmap.2(m,dendrogram='row',Rowv=T,Colv=T,trace='none',col=hm.colors,breaks=breaks,
          labRow = F, labCol = F,margins = c(2, 2),
          xlab="Lipid Correlation")

#FULL matrix
# mydataFull <- t(mydata)
# mydataFull[lower.tri(mydataFull)]<-mydata[lower.tri(mydata)]
# heatmap.2(mat,trace='none',col=col)

```

***

# Lipid Association Analysis

## Background
Sanity checks with the cleaned filtered data prior to performing downstream analysis  
- Clean data contains 416 individuals x 612 lipidome + 102 meta/phenodata
- These 416 inviduals have corresponding lipid and meta/cardiogram data   
- Raw lipid values were log2 transformed

## Read the data
```{r, fold.output=TRUE, fold.plot=FALSE}
dat <- read.csv("clean_dataset_LipidMetaCombined.csv", stringsAsFactors = FALSE, check.names = FALSE)
lipid <- dat[,102:ncol(dat)]
lipid$batch <- lipid$batch <- gsub(".*/", "", dat$date_enrol)
lipid_16 <- lipid[which(lipid$batch == "16"),]
lipid_17 <- lipid[which(lipid$batch == "17"),]
```

## Distribution of the lipidome data
Density distribution plots show that the lipid measures are normally distriuted  
Also, the distribution of lipids in both batches (year 2016, 2017) overlap  
Few, example (oxCE (18:2) [+2O]) show bimodal distribution

```{r density plots, fold.output=TRUE, fold.plot=FALSE}
pdf("Densityplot_log2_lipidome.pdf")
par(mfrow=c(3,4))
lipid_names <- colnames(lipid)[1:612]
for (i in 1:length(lipid_names)){
  lim=range(lipid[,i])
  a <- density(lipid_16[,i])
  b <- density(lipid_17[,i])
  max <- max(max(a$y), max(b$y))
  plot(density(lipid_16[,i], na.rm=TRUE), col="red", lwd=2, xlim=c(lim[1] - lim[1]*0.02,lim[2] + lim[2]*0.02), main=lipid_names[i], cex.main=0.8, ylim=c(0,max), xlab="Log2_lipid")
  lines(density(lipid_17[,i], na.rm=TRUE), col="blue", lwd=2)
  legend("topright", c("16", "17"), col=c("red", "blue"), lty=1, cex=0.5)
}
whatever <- dev.off()
```
[Link to a pdf version of the plot](Densityplot_log2_lipidome.pdf)  
**Few, example (oxCE (18:2) [+2O]), show bimodal distribution**

```{r, fig.width=3, fig.height=3,fig.show='hold',fig.align='left', fold.plot=FALSE}
lim=range(lipid$`oxCE (18:2) [+2O]`)
plot(density(lipid_16$`oxCE (18:2) [+2O]`, na.rm=TRUE), col="red", lwd=2, xlim=c(lim[1] - lim[1]*0.01,lim[2] + lim[2]*0.01), main="oxCE (18:2) [+2O]", cex.main=0.8,xlab="Log2_lipid")
lines(density(lipid_17$`oxCE (18:2) [+2O]`, na.rm=TRUE), col="blue", lwd=2)
```

## Lipid - Covariate Associations
Linear regression was used to explore the association between each lipid species and covariate.  
The P-values of association adjusted using the Benjamini–Hochberg procedure.   
covariates tested were: Age, Sex, HR, Diabetes status, Famhx_Heart, Blood Pressure (SBP and DBP), BMI, W/H ratio, CVDRisk (5 and 10 yr), batch
Note: for famhx_heart and diabetes - had individuals with either 0 (No); 1 (Yes); 11 (Unsure); the "Unsure" was changed to NA

```{r, fold.output=FALSE, fold.plot=FALSE}
cov <- dat[,c("sno", "sex", "age_calc", "HR", "BMIcalc", "waist_hip_ratio", "AvrgeSBP", "AvrgeDBP", "CVDRisk_10_yr","CVD_Risk5_yr", "famhx_heart", "diabetes")]
cov$batch <- lipid$batch
cov$diabetes[cov$diabetes == 11] <- NA
cov$famhx_heart[cov$famhx_heart == 11] <- NA

combined <- data.frame(matrix(ncol = 0, nrow = 612))
for (i in 2:ncol(cov)){
cov_name <- colnames(cov)[i]
d <- data.frame()
    for (j in 1:612){
fit <- summary(lm(lipid[,j] ~cov[,i]))
d[j,1] <- fit$coefficients[2,1]
d[j,2] <- fit$coefficients[2,4]
    }
colnames(d) <- c("beta", "pval")
d$fdr <- p.adjust(d$pval, method="BH")
colnames(d) <- paste(cov_name, colnames(d), sep="_")
rownames(d) <- colnames(lipid)[1:612]
combined <- cbind(combined, d)
}
write.csv(combined, "Assoc_Lipids_Covariates.csv", quote = FALSE, row.names=TRUE)
```
[Link to a csv file with the association results: **"Assoc_Lipids_Covariates.csv"** ](Assoc_Lipids_Covariates.csv)  
```{r, fold.output=FALSE, fold.plot=TRUE}
combined1 <- combined[,1:4]
combined1$sex_beta <- round(combined1$sex_beta, 2)
combined1$sex_pval <- formatC(combined1$sex_pval, format = "e", digits = 2)
combined1$sex_fdr <- formatC(combined1$sex_fdr, format = "e", digits = 2)
rownames(combined1) <- NULL
head(combined1[1:5, 1:3])

```

### Summary Lipid - Covariate Associations
The table summarises the number of lipids significantly (FDR < 0.05) associated with each covariate. The top lipid (lowest FDR) associated with each covariate is given together the beta estimate and FDR value. 
```{r, fold.output=FALSE, fold.plot=FALSE}

covlipid_summ <- data.frame()
for(i in 2:ncol(cov)){
names <- c(paste0(colnames(cov)[i], "_beta"), paste0(colnames(cov)[i], "_fdr"))
sub  <- combined[, names]
l <- length(which(sub[,2] < 0.05))
sub1 <- sub[order(sub[,2]),]
sub2 <- cbind(l, sub1[1,])
sub2$top <- rownames(sub2)
sub2$cov <- colnames(cov)[i]
sub3 <- sub2[, c(5,1,4,2,3)]
colnames(sub3) <- c("Covariate", "Tot_SigLipids", "TopLipid", "TopLipid_Beta", "TopLipid_FDR")
covlipid_summ <- rbind(covlipid_summ, sub3)
covlipid_summ$TopLipid_Beta <- round(covlipid_summ$TopLipid_Beta, 2)
covlipid_summ$TopLipid_FDR <- as.numeric(as.character(covlipid_summ$TopLipid_FDR))
covlipid_summ$TopLipid_FDR <- formatC(covlipid_summ$TopLipid_FDR, format = "e", digits = 2)
rownames(covlipid_summ) <- NULL
}
```

```{r}
library(kableExtra)
kable(covlipid_summ) %>%
kable_styling(bootstrap_options = c("striped", "hover"), font_size = 15, full_width = TRUE, position = "left") %>% 
footnote("oxCE (18:2) [+2O] is the top lipid associated with batch, this is the same lipid which has a strong bimodal distribition - with each peak representing the two year batches",  footnote_as_chunk = T, title_format = c("italic", "underline"), fixed_small_size = TRUE)
```

## Lipid - Echo Associations (no covariates)
Linear regression was used to explore the association between each lipid species and echo marker.  
The P-values of association adjusted using the Benjamini–Hochberg procedure.   
echo markers analysed are GLS_Peak, Eseptal, ELat, eavg, Eemed, EELat, EEavg, LAVavg, LA_Index, LVOTd, LVIDD, LVmass1, LVMass_Index

```{r, fold.output=FALSE, fold.plot=FALSE}
echo_sub <- dat[,c("sno", "GLS_Peak", "Eseptal", "ELat", "eavg", "Eemed", "EELat", "EEavg", "LAVavg", "LA_Index", "LVOTd", "LVIDD", "LVmass1", "LVMass_Index")]
echo_sub1 <- echo_sub[, -1]
echo <- log2(echo_sub1)
combined_echo <- data.frame(matrix(ncol = 0, nrow = 612))
for (i in 1:ncol(echo)){
echo_name <- colnames(echo)[i]
d_echo <- data.frame()
    for (j in 1:612){
fit_echo <- summary(lm(lipid[,j] ~echo[,i]))
d_echo[j,1] <- fit_echo$coefficients[2,1]
d_echo[j,2] <- fit_echo$coefficients[2,4]
    }
colnames(d_echo) <- c("beta", "pval")
d_echo$fdr <- p.adjust(d_echo$pval, method="BH")
colnames(d_echo) <- paste(echo_name, colnames(d_echo), sep="_")
rownames(d_echo) <- colnames(lipid)[1:612]
combined_echo <- cbind(combined_echo, d_echo)

}
write.csv(combined_echo, "Assoc_Lipids_Echomarkers_NoAdjCovariates.csv", quote = FALSE, row.names=TRUE)
```
[Link to a csv file with the association results: **"Assoc_Lipids_Echomarkers_NoAdjCovariates.csv"**](Assoc_Lipids_Echomarkers_NoAdjCovariates.csv)  

### Summary Lipid - Echo Associations (no covariates) 
The table summarises the number of lipids significantly (FDR < 0.05) associated with each echo marker.  
The top lipid (lowest FDR) associated with each echo marker is given together the beta estimate and FDR value

```{r, fold.output=FALSE, fold.plot=FALSE}
echolipid_summary <- data.frame()
for(i in 1:ncol(echo)){
names <- c(paste0(colnames(echo)[i], "_beta"), paste0(colnames(echo)[i], "_fdr"))
sub  <- combined_echo[, names]
l <- length(which(sub[,2] < 0.05))
sub1 <- sub[order(sub[,2]),]
sub2 <- cbind(l, sub1[1,])
sub2$top <- rownames(sub2)
sub2$echo <- colnames(echo)[i]
sub3 <- sub2[, c(5,1,4,2,3)]
colnames(sub3) <- c("echo_Marker", "TotSigLipids_NoAdjCov", "TopLipid_NoAdjCov", "TopLipid_Beta", "TopLipid_FDR")
echolipid_summary <- rbind(echolipid_summary, sub3)
}

echolipid_summary <- data.frame(echolipid_summary)
echolipid_summary$TopLipid_FDR <- formatC(echolipid_summary$TopLipid_FDR, format = "e", digits = 2)
echolipid_summary$TopLipid_Beta <- round(echolipid_summary$TopLipid_Beta, 2)
rownames(echolipid_summary) <- NULL
```

```{r}
library(kableExtra)
kable(echolipid_summary) %>%
kable_styling(bootstrap_options = c("striped", "hover"), font_size = 15, full_width = TRUE, position = "left")
```

## Echo - Batch Associations

There seems to be a difference between the distribution of the echo markers between the two collection batches  
The table shows the association results between each echo marker and batch   
The boxplot shows the distrbution of the log2 values of the each markers across the two batches (as an examples chose LVMass_Index, Eemed, EELat, Eseptal)

```{r, fold.output=FALSE, fold.plot=FALSE}
ec_dat <- data.frame()
for (i in 1:ncol(echo)){
echo_name <- colnames(echo)[i]
ec <- summary(lm(echo[,i] ~ cov$batch))
ec1 <- ec$coefficients[2,]
ec_dat <- rbind(ec_dat, ec1)
}
rownames(ec_dat) <- colnames(echo)
ec_dat1 <- ec_dat[, c(1, 4)]
colnames(ec_dat1) <- c("Beta", "Pval")
ec_dat1$Beta <- round(ec_dat1$Beta, 2)
ec_dat1$Pval <- formatC(ec_dat1$Pval, format = "e", digits = 2)
```

```{r}
library(kableExtra)
kable(ec_dat1) %>%
kable_styling(bootstrap_options = c("striped", "hover"), font_size = 15, full_width = TRUE, position = "left")
```


```{r dodge-st, fig.show = "hold", out.width = "40%"}
boxplot(echo$LVMass_Index ~ cov$batch, main= "LVMass_Index",xlab="Batch", ylab="Log2 echo marker", col=c("gold","lightgreen"))

boxplot(echo$Eemed ~ cov$batch, main= "Eemed",xlab="Batch", ylab="Log2 echo marker", col=c("gold","lightgreen"))

boxplot(echo$EELat ~ cov$batch, main= "EELat",xlab="Batch", ylab="Log2 echo marker", col=c("gold","lightgreen"))

boxplot(echo$Eseptal ~ cov$batch, main= "Eseptal",xlab="Batch", ylab="Log2 echo marker", col=c("gold","lightgreen"))

```

## Lipid - Echo Associations (with covariates)
Multiple Linear regression was used to explore the association between each lipid species and echo marker, while adjusting for covariates in the model.  
The P-values of association adjusted using the Benjamini–Hochberg procedure.   
**Echo markers** analysed are: *GLS_Peak, Eseptal, ELat, eavg, Eemed, EELat, EEavg, LAVavg, LA_Index, LVOTd, LVIDD, LVmass1, LVMass_Index*  
**Covariates adjusted for** are: *Age, Sex, HR, Diabetes status, Famhx_Heart, Blood Pressure (SBP and DBP), BMI, W/H ratio, CVDRisk (5 and 10 yr), batch* 

```{r, fold.output=FALSE, fold.plot=FALSE}
c <- cov
combined_echo_adjCov <- data.frame(matrix(ncol = 0, nrow = 612))
for (i in 1:ncol(echo)){
echo_name <- colnames(echo)[i]
echo_adjCov <- data.frame()
    for (j in 1:612){
fit_echo_adjcov <- summary(lm(lipid[,j] ~echo[,i] + c$sex + c$age_calc + c$HR  + c$BMIcalc + c$waist_hip_ratio + c$AvrgeSBP + c$AvrgeDBP + c$CVDRisk_10_yr + c$CVD_Risk5_yr + c$famhx_heart + c$diabetes + c$batch))
echo_adjCov[j,1] <- fit_echo_adjcov$coefficients[2,1]
echo_adjCov[j,2] <- fit_echo_adjcov$coefficients[2,4]
    }
colnames(echo_adjCov) <- c("beta", "pval")
echo_adjCov$fdr <- p.adjust(echo_adjCov$pval, method="BH")
colnames(echo_adjCov) <- paste(echo_name, colnames(echo_adjCov), sep="_")
rownames(echo_adjCov) <- colnames(lipid)[1:612]
combined_echo_adjCov <- cbind(combined_echo_adjCov, echo_adjCov)
}
write.csv(combined_echo_adjCov, "Assoc_Lipids_Echo_AdjustedForCovariates.csv", quote = FALSE, row.names=TRUE)
```
[Link to a csv file with the association results: **"Assoc_Lipids_Echo_AdjustedForCovariates.csv"**](Assoc_Lipids_Echo_AdjustedForCovariates.csv)  

### Summary Lipid - Echo Associations (with covariates)
The table summarises the number of lipids significantly (FDR < 0.05) associated with each echo marker.  
The top lipid (lowest FDR) associated with each echo marker (adjusted for covariates) is given together the beta estimate and FDR value

```{r, fold.output=FALSE, fold.plot=FALSE}
echolipidAdjCov_summary <- data.frame()
for(i in 1:ncol(echo)){
names <- c(paste0(colnames(echo)[i], "_beta"), paste0(colnames(echo)[i], "_fdr"))
sub  <- combined_echo_adjCov[, names]
l <- length(which(sub[,2] < 0.05))
sub1 <- sub[order(sub[,2]),]
sub2 <- cbind(l, sub1[1,])
sub2$top <- rownames(sub2)
sub2$echo<- colnames(echo)[i]
sub3 <- sub2[, c(5,1,4,2,3)]
colnames(sub3) <- c("echo_Marker", "TotSigLipids_AdjCov", "TopLipid_AdjCov", "TopLipid_Beta", "TopLipid_FDR")
echolipidAdjCov_summary  <- rbind(echolipidAdjCov_summary, sub3)
}

echolipidAdjCov_summary <- data.frame(echolipidAdjCov_summary)
echolipidAdjCov_summary$TopLipid_FDR <- formatC(echolipidAdjCov_summary$TopLipid_FDR, format = "e", digits = 2)
echolipidAdjCov_summary$TopLipid_Beta <- round(echolipidAdjCov_summary$TopLipid_Beta, 2)
rownames(echolipidAdjCov_summary) <- NULL
```

```{r}
library(kableExtra)
kable(echolipidAdjCov_summary) %>%
kable_styling(bootstrap_options = c("striped", "hover"), font_size = 15, full_width = TRUE, position = "left")
```


