---
title: "UKBiobank"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
    theme: cosmo
    df_print: paged
    highlight: tango
    # code_folding: hide
# output:
#   epuRate::BAKER:
#     toc: TRUE
#     number_sections: FALSE
#     code_folding: "show"
---

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });

});
</script>

***

# Play around with UKBiobank

## Plots for Mike presentation

### Load Libraries

```{r warning=FALSE, results='hide'}
library(summarytools)
library(stringr)
st_options(plain.ascii = FALSE,        # Always use this option in Rmd documents
           style        = "rmarkdown", # Always use this option in Rmd documents
           footnote     = NA,          # Makes html-rendered results more concise
           subtitle.emphasis = FALSE)  # Improves layout with some rmardown themes
library(lobstr)
library(knitr)
opts_chunk$set(results = 'asis',      # Can also be set at the chunk-level
               comment = NA,
               prompt  = FALSE,
               cache   = FALSE)
library(ukbtools)
library(tidyverse)
```

### Load data

Load data in the cluster and save selected fields for reports

```{r}
# # to run in the cluster
# library(stringr)
# # load("/baker/datasets/ukb55469/R-env/my_ukb_data.rda")
# nm <- names(my_ukb_data)
# field.patterns.illness <- tolower(c("eid","cholesterol", "DL_", "Vitamin_d", "Calcium","Creatinine","IGF1",
#                                     "Creactive","Lipoprotein","lbumin","noncancer_illness_code_selfreported_f20002"))
# sel.fields.illness <- unlist(sapply(field.patterns.illness,function(x){nm[str_detect(nm,x)]}))
# names(sel.fields.illness) <- NULL
# mydata.illness <- my_ukb_data[,sel.fields.illness]
# save(mydata.illness,file="~/my_ukb_data_selectedfields_illness.rda")
# 
# #########
# field.patterns <- tolower(c("eid","cholesterol", "DL_", "Vitamin_d", "Calcium","Creatinine","IGF1","Creactive","Lipoprotein","lbumin"))
# sel.fields <- unlist(sapply(field.patterns,function(x){nm[str_detect(nm,x)]}))
# names(sel.fields) <- NULL
# mydata <- my_ukb_data[,sel.fields]
# save(mydata,file="~/my_ukb_data_selectedfields.rda")
```

*** 

### Load files from cluster

Files from cluster are mounted in sgcluster_workspace in my laptop
Load Rodrigo's functions too

```{r}
# load("/sysgen/workspace/users/sruizcarmona/WORK/BIOINFO/UKBIOBANK/r-objects/my_ukb_data_selectedfields_illness.rda")
# load("/sysgen/workspace/users/sruizcarmona/WORK/BIOINFO/UKBIOBANK/r-objects/my_ukb_data_selectedfields.rda")
load("~/sgcluster/WORKSPACE/WORK/BIOINFO/UKBIOBANK/RDA_FILES/my_ukb_data_selectedfields_illness.rda")
load("~/sgcluster/WORKSPACE/WORK/BIOINFO/UKBIOBANK/RDA_FILES/my_ukb_data_selectedfields.rda")
source("../ukb_functions.R")
```

***

## Data reports

Summarize a couple of variables

```{r}
sum.fields <- c("cholesterol_f30690_0_0","cholesterol_f30690_1_0","cholesterol_assay_date_f30691_0_0","cholesterol_assay_date_f30691_1_0")
dfSummary(mydata.illness[,sum.fields],style="grid",headings=F,valid.col=T,tmp.img.dir = "/tmp",varnumbers = T)
```

Extract info from selfreported diseases

* Heart attack / miocardial infacrtion
* Type 2 Diabetes

```{r}
# self reported subsets
# heart attack / miocardial 1075
heartattack <- extract_self_reported_nc(mydata.illness,1075)
# t2d 1223
t2d <- extract_self_reported_nc(mydata.illness,1223)
```

### Heart Attack

Heart attack vs high ldl

```{r}
# ldl in mmol/L
# high ldl is above 130 mg/dL, which corresponds to 3.4 mmol/L
# https://medlineplus.gov/ldlthebadcholesterol.html
mydata.illness$heartattack <- "No"
mydata.illness$heartattack[mydata.illness$eid %in% heartattack$eid] <- "Yes"
chol <- mydata.illness[!is.na(mydata.illness$ldl_direct_f30780_0_0),] %>% 
  mutate(highldl = ldl_direct_f30780_0_0 > 3.4)
# dfSummary(mydata.illness[,sum.fields],style="grid",headings=F,valid.col=F,tmp.img.dir = "/tmp")
print(ctable(x = chol$heartattack, y = chol$highldl , prop = "r",style="grid",headings=F,plain.ascii=T,justify="c"),
      method = "render")
```

## Reports with small set

The idea is to check with the small subset and then use the session in the cluster to create the same reports for the whole dataset

```{r}
# load("~/sgcluster_home/small_ukb_set.rda")
```


```{r}
# sum.patterns <- c("ethnic_background")
# sum.patterns <- c("4272")
# sum.fields<- unlist(sapply(sum.patterns,function(x){names(small_ukb_set)[str_detect(names(small_ukb_set),x)]}))
# sum.fields
# dfSummary(small_ukb_set[,sum.fields],style="grid",headings=F,valid.col=T,tmp.img.dir = "/tmp",varnumbers = T)
```
```{r}
# load("~/sgcluster/eth.rda")
```


```{r}
# eth <- as.data.frame(k)
# colnames(eth)<-'Individuals'
# eth$Ethnicity <- rownames(eth)
# # require(scales)
# options(scipen=10000)
# point <- format_format(big.mark = " ", decimal.mark = ",", scientific = FALSE)
# ggplot(eth, aes(x=Ethnicity,y=Individuals)) +
#   geom_bar(stat='identity') +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 45,hjust = 1))

```

***

# Tidy data for visits and medications

```{r}
load("../../RDA_FILES/preprocessed_rdas/small_ukb_set.rda")
```

## Clean small dataset

* Test with date of assessment (4 possible fields)
* And first medication reported for each of the visits

```{r}
# k <- small_ukb_set[,c('eid','date_of_attending_assessment_centre_f53_0_0','date_of_attending_assessment_centre_f53_1_0','date_of_attending_assessment_centre_f53_2_0','date_of_attending_assessment_centre_f53_3_0','treatmentmedication_code_f20003_0_0','treatmentmedication_code_f20003_1_0','treatmentmedication_code_f20003_2_0','treatmentmedication_code_f20003_3_0')]
date_fields <- str_subset(names(small_ukb_set),"date_of_attending_assessment")
med_fields <- str_subset(names(small_ukb_set),"treatmentmedication_code_")
sel_fields <- c('eid',date_fields,med_fields)
# k <- small_ukb_set[,c('eid','date_of_attending_assessment_centre_f53_0_0','date_of_attending_assessment_centre_f53_1_0','date_of_attending_assessment_centre_f53_2_0','date_of_attending_assessment_centre_f53_3_0','treatmentmedication_code_f20003_0_0','treatmentmedication_code_f20003_0_1')]
k <- small_ukb_set[,sel_fields]
names(k) <- str_replace(names(k),"treatmentmedication_code_f20003_","med_")
names(k) <- str_replace(names(k),"date_of_attending_assessment_centre_f53_","visit_")
k$med_2_0 <- c(rep(NA,5),"12345",rep(NA,4))
names(k)[names(k) %in% "visit"] <- str_replace(names(k),"_0$","")
# testdf <- k[1:2,]
```

```{r}
library(lubridate)
```

```{r}
# as.data.frame(k %>% pivot_longer(cols=c(names(k)[grepl("visit",names(k))]), names_to="visit", values_to = "date") %>% separate(visit,into=c("k","visit_n"),convert=T,sep="_") %>% select(-k))
k %>%
  # mutate_if(is.numeric, as.character) %>% 
  # mutate_if(is.Date,as.character) %>% 
  # pivot_longer(-eid,values_drop_na = T) %>% 
  # # mutate(visit=case_when(grepl("visit",name) ~ str_replace(name,"_0$",""))) %>% 
  # separate(name,into=c("type","visit_n","array_n"),sep="_",convert=T) %>%
  # mutate(visit_n=visit_n+1,array_n=paste0("med_",array_n+1)) %>%
  # pivot_wider(id_cols = visit_n, values_from = value, names_from = type) %>% 
  data.frame()
```

```{r}
coalesce_by_column <- function(df) {
  return(coalesce(df[1], df[2]))
}
# https://stackoverflow.com/questions/45515218/combine-rows-in-data-frame-containing-na-to-make-complete-row
```


```{r}
# as.data.frame(k %>% pivot_longer(cols=c(names(k)[grepl("visit",names(k))]), names_to="visit", values_to = "date") %>% separate(visit,into=c("k","visit_n"),convert=T,sep="_") %>% select(-k))
k %>%
  mutate_if(is.numeric, as.character) %>% 
  mutate_if(is.Date,as.character) %>% 
  pivot_longer(-eid,values_drop_na = T) %>% 
  separate(name,into=c("type","visit_n","array_n"),sep="_",convert=T) %>% 
  mutate(visit_n=visit_n+1,array_n=paste0("med_",array_n+1)) %>%
  pivot_wider(values_from = value, names_from = type)  %>%
  mutate(date=visit) %>%
  select(-visit) %>%
  pivot_wider(names_from = array_n,values_from = med) %>%
  group_by(eid,visit_n) %>%
  summarise_all(coalesce_by_column) %>%
  ungroup() %>%
  data.frame()
```

```{r}
k %>%
  mutate_if(is.numeric, as.character) %>% 
  mutate_if(is.Date,as.character) %>% 
  pivot_longer(-eid,values_drop_na = T) %>% 
  separate(name,into=c("type","visit_n","array_n"),sep="_",convert=T) %>% 
  mutate(visit_n=visit_n+1,array_n=paste0("med_",array_n+1)) %>%
  pivot_wider(values_from = value, names_from = type)  %>%
  mutate(date=as.Date(visit)) %>%
  select(-visit) %>%
  pivot_wider(names_from = array_n,values_from = med) %>%
  group_by(eid,visit_n) %>%
  summarise_all(coalesce_by_column) %>%
  ungroup() %>%
  pivot_longer(starts_with('med'), names_to="med_n",values_to='med',values_drop_na = T) %>%
  filter(med %in% c('1140868226','2038460150')) %>%
  pivot_wider(names_from=med_n,values_from=med) %>%
  data.frame()
```

```{r}
mydf <- data.frame(matrix(c(1, "a", "b", "c",  "1", "2", "3",2,"d", "e", "f",  "4", '5', "6"),byrow =T,ncol=7))
names(mydf) <- c("eid", "i1", "i2", "i3", "d1", "d2", "d3")
mydf %>% 
  pivot_longer(starts_with("i"),names_to = 'k1',values_to = "i") %>% 
  separate(k1,into = c("val1","id1"),sep="i") %>% 
  pivot_longer(starts_with("d"),names_to = 'k2',values_to = "d") %>%
  separate(k2,into = c("val2","id2"),sep="d") %>%
  filter(id1==id2) %>%
  select(eid,i,d) %>%
  data.frame()

```

example with dates and icd10

```{r}
load("/baker/datasets/ukb55469/phenotype_traits/R-env/preprocessed_rdas/small_ukb_set.rda")
date_fields <- str_subset(names(small_ukb_set),"date_of_first_inpatient_diagnosis_icd10_f41280")
med_fields <- str_subset(names(small_ukb_set),"diagnoses_icd10_f41270")
sel_fields <- c('eid',date_fields,med_fields)
k <- small_ukb_set[,sel_fields]
names(k) <- str_replace(names(k),"date_of_first_inpatient_diagnosis_icd10_f41280_","icd10diagdate_")
names(k) <- str_replace(names(k),"diagnoses_icd10_f41270_","icd10_")
```

```{r}
k[1,] %>%
  select_if(~sum(!is.na(.)) > 0)%>% 
  data.frame()
```


```{r}
k %>%
  mutate_if(is.numeric, as.character) %>% 
  mutate_if(is.Date,as.character) %>% 
  pivot_longer(-eid,values_drop_na = T) %>% 
  separate(name,into=c("type","visit_n","array_n"),sep="_",convert=T) %>%
  mutate(visit_n=visit_n+1,array_n=paste0("icd10_",array_n+1)) %>%
  pivot_wider(values_from = value, names_from = type)  %>%
  mutate(date=icd10diagdate) %>%
  select(eid,date,icd10) %>%
  # pivot_wider(names_from = array_n,values_from = icd10) %>%
  # group_by(eid,visit_n) %>%
  # summarise_all(coalesce_by_column) %>%
  # ungroup() %>%
  data.frame()
```

